---
title: "Figures for Manuscript"
author: "Linh Dang"
editor: visual
execute:
  warning: false
  message: false
format:
  html:
    code-fold: true
    toc: true
    toc-location: left
    toc-depth: 3
    fig-responsive: true
    fig-lightbox: true
---

## Introduction

This is a script to recreate figures for the manuscript.

**Library**

```{r}
#| message: false

library(phyloseq)
library(reshape2)
library(microViz)
library(microbiome)
library(vegan)
library(lme4)       # Linear Mixed Models (for repeated measures)
library(lmerTest)   # P-values for Mixed Models
library(emmeans)    # Pairwise comparisons
library(Wrench)
library(rstatix)
library(ggpubr)
library(tidyverse)
library(VennDiagram)
library(ComplexUpset)
library(patchwork)
library(pairwiseAdonis)
library(MASS)
rm(list = ls())
```

## Chapter 1

**Load data**

```{r}
## remove old object
rm(list = setdiff(ls(), c(lsf.str(), "Kolors", "Kolors_Chap2")))
## load objects
pseq <- dirname(rstudioapi::getSourceEditorContext()$path) %>% 
  file.path("data/longitudinal_contaminant_survey/NCT_2026April_v0.rds") %>% 
  readRDS()
N <- 500
pseq <- pseq %>% 
  #ps_filter(Control_type %in% c("Buffer", "Paraffin", "PCR", "Sequencing")) %>%
  ps_filter(Control_type %in% c("Buffer", "Paraffin", "PCR")) %>% 
  prune_samples(sample_sums(.) >=N, .) %>% 
  ps_get() %>% 
  microViz::tax_filter(min_prevalence = 1,
                       prev_detection_threshold = 2,
                       min_total_abundance = 1e-6)
## add year and quarter to metadata
pseq <- pseq %>% 
  ps_mutate(Year = as.character(lubridate::year(Seq_date))) %>% 
  ps_mutate(Quarter = as.character(lubridate::quarter(Seq_date)))
```

------------------------------------------------------------------------

### 1.B

```{r}
#| fig-width: 14
#| fig-height: 9
#| fig-lightbox: true
require(patchwork)
#lst <- pseq %>% 
#  AlphaPlotWrapper_Violin(strata = "Control_type", x_label = "Control type")
#plt <- lst$Observed + lst$Shannon + lst$InvSimpson + plot_layout(guides = "collect") +
main_v <- "Control_type"
covariate_v <- c("Person", "Year", "Quarter")
#covariate_v <- NULL
sample_data(pseq)$Control_type <- factor(sample_data(pseq)$Control_type, levels = c("Buffer", "Paraffin", "PCR"))
res1 <- AlphaPlot_Violin_glm(PhyloObj = pseq, index = "Observed", main_var = main_v, covariate_var = covariate_v, x_label = "Control type", y_label = "Number of observed species", add_legend = T, gg_title = NULL, gg_subtitle = NULL, m_Kolors = Kolors, m_family = NULL, is_nb = TRUE)
res2 <- AlphaPlot_Violin_glm(PhyloObj = pseq, index = "Shannon", main_var = main_v, covariate_var = covariate_v, x_label = "Control type", y_label = "Shannon index", add_legend = T, gg_title = NULL, gg_subtitle = NULL, m_Kolors = Kolors, m_family = Gamma(link = "inverse"), is_nb = FALSE)

#plt <- lst$Observed + lst$Shannon + plot_layout(guides = "collect") +
plt <- res1$plt + res2$plt + plot_layout(guides = "collect")+
  plot_annotation(title = "Alpha diversity stratified by control types") &
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5))
## save plt to png
#ggsave("./Figures_for_Manuscript_files/saved_png/1.B.svg", plt, width = 8, height = 6, dpi = 300, limitsize = F)
plt

```

```{r}
## pairwise comparison
df <- summary(res2$const, infer = c(TRUE, TRUE)) %>% 
  as.data.frame() %>% 
  mutate(exp = exp(estimate),
         LCL = exp(lower.CL),
         UCL = exp(upper.CL))
summary(res2$const, infer = c(TRUE, TRUE))
```

**Figure 1.2** Alpha diversity of NCT samples by control type: observed species richness, Shannon index, and inverse Simpson index (left to right respectively). Statistically significant pairwise differences are indicated.

------------------------------------------------------------------------

### 1.C

PLAYGROUND

```{r}
## for beta analyses dbrda()
Y <- abundances(pseq) %>% t()
#Y <- decostand(Y, method = "hellinger")
bray_dist = vegan::vegdist(Y, method = "bray")
df_meta <- meta(pseq)
df_meta$Control_type <- factor(df_meta$Control_type)
df_meta$Person <- factor(df_meta$Person)
df_meta$Year <- factor(df_meta$Year)
df_meta$Quarter <- factor(df_meta$Quarter)

#dbrda_mod <- dbrda(bray_dist ~ Control_type + Condition(Person + Year + Quarter), data = df_meta)
dbrda_mod <- dbrda(bray_dist ~ Control_type + Person + Year + Quarter, data = df_meta)
results <- anova(dbrda_mod, by = "margin", permutations = 9999) 
%>% as.data.frame()
results <- results %>% 
  rownames_to_column(var = "Variable")
library(effectsize)

# ## plot
# plot(dbrda_mod, type = "n")
# points(dbrda_mod, display = "sites", pch = 21,
#        col = "black",
#        bg = as.numeric(factor(df_meta$Control_type)))
# legend("topright",
#        legend = levels(factor(df_meta$Control_type)),
#        pch = 21, pt.bg = 1:length(levels(factor(df_meta$Control_type))))
# ##
# plot(dbrda_mod, type = "n")
# points(dbrda_mod, display = "sites", pch = 21,
#        col = "black",
#        bg = as.numeric(factor(df_meta$Control_type)))
# text(dbrda_mod, display = "bp", col = "blue")  # biplot for predictors

sites <- scores(dbrda_mod, display = "sites", scaling = 2)
sites <- as.data.frame(sites)
sites$Control_type <- df_meta$Control_type

p <- ggplot(sites, aes(x = dbRDA1, y = dbRDA2, colour = Control_type, fill = Control_type)) +
  geom_point(size = 3, alpha = 0.8) +
  stat_ellipse(level = 0.95, geom = "polygon", alpha = 0.2, colour = NA) +
  labs(
    title = "dbRDA of microbiome beta diversity",
    subtitle = "Constrained by Control_type + LP + year + season",
    x = paste0("dbRDA1 (", round(summary(dbrda_mod)$constr.proportion[1] * 100, 1), "%)"),
    y = paste0("dbRDA2 (", round(summary(dbrda_mod)$constr.proportion[2] * 100, 1), "%)")
  ) +
  theme_bw() +
  theme(
    legend.position = "right"
  )
p
```

```{r}
#| fig-width: 18
#| fig-height: 7
#| fig-lightbox: true
require(patchwork)
plt_wrench <- pseq %>% 
  WrenchWrapper(grp = "Control_type") %>% 
  BetaPlot(strata_f = "Control_type", dis_method = "bray", ordination_method = "PCoA", title_method = "NCT samples stratified by control types \n Wrench normalization")
ggsave("./Figures_for_Manuscript_files/saved_png/1.C.svg", plt_wrench, width = 8, height = 6, dpi = 300)
##------------------------
plt_rar <- pseq %>% 
  rarefy_even_depth() %>% 
  BetaPlot(strata_f = "Control_type", dis_method = "bray", ordination_method = "PCoA", title_method = "NCT samples stratified by control types \n Rarefaction")
#plt <- plt_wrench + plot_layout(guides = "collect")
plt <- plt_wrench + plt_rar + plot_layout(guides = "collect")
## save to png
ggsave("./Figures_for_Manuscript_files/saved_png/S1_BetaDiversity.svg", plot = plt, width = 10, height = 6, dpi = 300)
plt
```

**Figure 1C**: Beta diversity ordination of NCT samples using Wrench and rarefaction as normalization (left to right). Group centroids and PERMANOVA p-values are shown.

------------------------------------------------------------------------

Table for pairwise

```{r}
require(pairwiseAdonis)
bray_dist = phyloseq::distance(pseq, method="bray")
p_res <- pairwise.adonis2(x = Control_type, data = meta(pseq), strata = "Person", nperm = 9999)
#p_res <- pairwise.adonis2(x = bray_dist ~ Control_type, data = meta(pseq), nperm = 9999)
```

```{r}
### adonis2() with formular
#adonis2(formula = bray_dist ~ Person + Year + Quarter + Control_type, data = meta(pseq), permutations = 9999, by = "margin")

meta <- meta(pseq)
dist <- pseq %>% 
  WrenchWrapper(grp = "Control_type") %>% 
  #rarefy_even_depth() %>% 
  phyloseq::distance(method = "bray")
pairwise_levels <- combn(unique(meta$Control_type), 2, simplify = FALSE)

results <- lapply(pairwise_levels, function(x) {
  idx <- meta$Control_type %in% x
  meta_sub <- droplevels(meta[idx, ])
  dist_sub <- as.dist(as.matrix(dist)[idx, idx])

  fit <- adonis2(dist_sub ~ Person + Year + Quarter + Control_type,
                 data = meta_sub,
                 #method = "bray",
                 permutations = 9999,
                 by = "margin")
  
  data.frame(
    group1 = x[1],
    group2 = x[2],
    p_value = fit["Control_type", "Pr(>F)"],
    F_model = fit["Control_type", "F"],
    R2 = fit["Control_type", "R2"]
  )
})
print(results)
```

### 1.D

```{r}
#| fig-width: 10
#| fig-height: 7
#| fig-lightbox: true
library(svglite)
library(ComplexHeatmap)
plt <- HeatMap_MicroViz(ps = pseq, rnk = "genus", tax_trans = "compositional", tax_order_by = sum, n_taxa = 15, my_title = "Top 15 genera by sum")

plt
```

**Figure 1D**: Heatmap of the top 15 most abundant taxa in NCT samples at genus level.

Get a list of most common species

```{r}
## number of genus 
pseq_tmp <- pseq %>% 
  microViz::tax_agg(rank = "genus")
df <- pseq_tmp %>% 
  taxa_sums() %>% 
  enframe(name = "Taxa", value = "Reads") %>% 
  arrange(desc(Reads))
df$CumPercent <- cumsum(df$Reads) / sum(df$Reads) * 100
df %>% head(n=30) %>% 
  write_csv(., file="tmp.txt")
df_prev <- pseq %>% prevalence(count = F) %>% enframe(name = "Taxa", value = "Prevalence")
total_reads <- pseq %>% taxa_sums() %>% sum()
df_abund <- pseq %>% taxa_sums() %>% enframe(name = "Taxa", value = "Abundance")
df <- df_prev %>% 
  left_join(., df_abund, by = "Taxa") %>% 
  mutate(Abundance = Abundance / total_reads) %>% 
  arrange(desc(Abundance)) %>% 
  filter(Abundance >=0.001)
```

------------------------------------------------------------------------

------------------------------------------------------------------------

## Chapter 2

------------------------------------------------------------------------

Load data

```{r}

## remove old object
rm(list = setdiff(ls(), c(lsf.str(), "Kolors", "Kolors_Chap2")))
pseq <- rprojroot::find_rstudio_root_file() %>% 
  file.path("data/Chap2/px_J_U_NCT_v02.rds") %>% 
  readRDS() %>% 
  ps_get() %>% 
  tax_filter(min_total_abundance = 2, min_prevalence = 2)
## edit meta
pseq <- pseq %>% 
  ps_mutate(LP = ifelse(person == "J", "LP1", "LP2")) %>% 
  ps_mutate(Environment = HOOD_BENCH) %>% 
  ps_select(-c("HOOD_BENCH", "person"))
```

### Alpha

```{r}
lst <- AlphaPlotWrapper_Violin(pseq,roundUp = TRUE,  strata = "LP", m_Kolors = Kolors_Chap2, x_label = "LP")
plt <- lst$Observed + lst$Shannon + lst$InvSimpson + plot_layout(guides = "collect")
ggsave("./Figures_for_Manuscript_files/saved_png/2.B.svg", plot = plt, width = 8, height = 6, dpi = 300)
plt

## Alternative
al1 <- AlphaPlot_Violin_glm(pseq, index = "Observed", main_var = "Environment", covariate_var = "LP", x_label = "Environment", y_label = "Number of observed species", add_legend = T, gg_title = NULL, gg_subtitle = NULL, m_Kolors = Kolors_Chap2[3:4], m_family = NULL, is_nb = T)
al2 <- AlphaPlot_Violin_glm(pseq, index = "Shannon", main_var = "Environment", covariate_var = "LP", x_label = "Environment", y_label = "Shannon index", add_legend = T, gg_title = NULL, gg_subtitle = NULL, m_Kolors = Kolors_Chap2[3:4], m_family = gaussian(), is_nb = F)

plt <- al1$plt + al2$plt + plot_layout(guides = "collect")
ggsave("./Figures_for_Manuscript_files/saved_png/S2.B.svg", plot = plt, width = 8, height = 6, dpi = 300)
  
al2$const %>% 
  summary(infer = c(TRUE, TRUE)) %>% 
  as.data.frame() %>% 
  mutate(exp = exp(estimate),
         LCL = exp(lower.CL),
         UCL = exp(upper.CL)) %>% 
  dplyr::select(contrast, estimate, exp, LCL, UCL, p.value) %>% 
  write_csv("tmp.txt")
```

**Figure 2B**: Alpha diversity of NCT samples stratified by persons who processed samples.

------------------------------------------------------------------------

```{r}
rich_meta <- merge(pseq %>% sample_data(), pseq %>% estimate_richness(), by = "row.names")
model_lm <- lm(Shannon ~ Environment, data = rich_meta)
summary(model_lm)
```

Environment after accounting for LP

```{r}
model_lmer <- lmer(Shannon ~ Environment + (1 | LP), data = rich_meta)
summary(model_lmer)
```

------------------------------------------------------------------------

PLAYGROUND

```{r}
df_data <- meta(pseq)
df_data$Shannon <- pseq %>% 
  #rarefy_even_depth() %>% 
  #WrenchWrapper(grp = "Control_type") %>% 
  otu_table() %>% 
  t() %>% 
  diversity(index = "shannon")
  #vegan::specnumber()
fit <- lm(Shannon ~ Environment, data = df_data)
emm <- emmeans(fit, ~ Environment)

# Pairwise comparisons
pairs(emm)

## 
require(pairwiseAdonis)
bray_dist = phyloseq::distance(pseq, method="bray")
p_res <- pairwise.adonis2(x = bray_dist ~ Environment, data = meta(pseq), nperm = 9999)
p_res2 <- pairwise.adonis2(x = bray_dist ~ Environment, data = meta(pseq), strata = "LP", nperm = 9999)
```

### Beta

```{r}
##| fig-width: 16
##| fig-height: 9
##| fig-lightbox: true
pseq_clr <- pseq %>% microbiome::transform(transform = "clr")
#dist_matrix <- phyloseq::distance(pseq_clr, method = "euclidean")
plt1 <- BetaPlot_v02(pseq_clr, strata_f = "Environment", color_f = "LP", shape_f = "Environment", ellipse_f = "Environment", dis_method = "euclidean", title_method = "Hood vs. Bench")
plt2 <- BetaPlot_v02(pseq_clr, strata_f = "LP", color_f = "LP", shape_f = "Environment", ellipse_f = "LP", dis_method = "euclidean", title_method = "Between Laboratory Personnels")
plt <- plt1 + plt2 + plot_layout(guides = "collect")
## save to png
ggsave("./Figures_for_Manuscript_files/saved_png/2.C.svg", plot = plt2, width = 8, height = 6, dpi = 300)
plt2
```

**Figure 2C**: Beta diversity cluster by environmental conditions (left), and technicians (right). Colors depict different technicians, while shapes depict different environmental conditions.

Consider LP with Environment as covariate

Method 1

```{r}
bray_dist <- pseq %>% 
  rarefy_even_depth() %>% 
  phyloseq::distance(method="bray")
clr_euclidean_dist <- pseq %>% 
  microbiome::transform(transform = "clr") %>% 
  phyloseq::distance(method="euclidean")
beta_model_TA <- adonis2(
  bray_dist ~ Environment + LP,
  data = df_data,
  permutations = 9999, 
  by = "terms"
)

beta_model_TA
```

Method 2

```{r}
df_meta <- meta(pseq)
beta_model_block <- adonis2(
  bray_dist ~ LP,
  data = df_meta,
  #permutations = 9999
  permutations = how(
    blocks = df_meta$Environment
  )
)
```

New approach: dbrda with anova

```{r}
Y <- pseq %>% 
  #microbiome::transform(transform = "clr") %>% 
  WrenchWrapper(grp = "LP") %>% 
  abundances() %>% 
  t()
#Y <- decostand(Y, method = "hellinger")
bray_dist = vegan::vegdist(Y, method = "bray") ## euclidean
df_meta <- meta(pseq)
#df_meta$block <- with(df_meta, paste(LP, Environment, sep = "_"))
df_meta$LP <- factor(df_meta$LP)
df_meta$Environment <- factor(df_meta$Environment)
#df_meta$block <- factor(df_meta$block)

#dbrda_mod <- dbrda(bray_dist ~ Control_type + Condition(Person + Year + Quarter), data = df_meta)
dbrda_mod <- dbrda(bray_dist ~ LP, data = df_meta)
results <- anova(dbrda_mod, by = "term", permutations = how(
    blocks = df_meta$Environment)) %>% as.data.frame()
results <- results %>% 
  rownames_to_column(var = "Variable")
```

```{r}
Y <- pseq %>% 
  #WrenchWrapper(grp = "LP") %>% 
  rarefy_even_depth(rngseed = 210488) %>% 
  #microbiome::transform(transform = "clr") %>% 
  abundances() %>% 
  t()
#Y <- decostand(Y, method = "hellinger")
bray_dist = vegan::vegdist(Y, method = "bray")
df_meta <- meta(pseq)
# df_meta$block <- with(df_meta, paste(LP, Environment, sep = "_"))
df_meta$LP <- factor(df_meta$LP)
df_meta$Environment <- factor(df_meta$Environment)
#df_meta$block <- factor(df_meta$block)

#dbrda_mod <- dbrda(bray_dist ~ Control_type + Condition(Person + Year + Quarter), data = df_meta)
dbrda_mod_1 <- dbrda(bray_dist ~ Environment, data = df_meta)
results_1 <- anova(dbrda_mod_1, by = "term", permutations = how(
    blocks = df_meta$LP)) %>% as.data.frame()
results_1 <- results_1 %>% 
  rownames_to_column(var = "Variable") %>% 
  mutate(R2 = SumOfSqs / sum(SumOfSqs)) %>% 
  filter(Variable == "Environment")

dbrda_mod_2 <- dbrda(bray_dist ~ LP, data = df_meta)
results_2 <- anova(dbrda_mod_2, by = "term", permutations = how(
    blocks = df_meta$Environment)) %>% as.data.frame()
results_2 <- results_2 %>% 
  rownames_to_column(var = "Variable") %>% 
  mutate(R2 = SumOfSqs / sum(SumOfSqs)) %>% 
  filter(Variable == "LP")
res <- rbind(results_1, results_2)
```

------------------------------------------------------------------------

------------------------------------------------------------------------

## Chapter 3

### 3.1 Assessment Using Longitudinal Negative Control Survey

Inspect NCT for chapter 3

```{r}
pseq_C3 <- rprojroot::find_rstudio_root_file() %>% 
  file.path("data/Chap3", "pseq_Proj5_postFilter_v04.rds") %>% 
  readRDS()
df_C3 <- meta(pseq_C3) %>% 
  filter(ffpe.bulk == "FFPE")
```

Load data

```{r}
rm(list = setdiff(ls(), c(lsf.str(), "Kolors", "Kolors_Chap2")))
## Load significant taxa in NCT survey
thres_prev_NCT <- 0.25 ## threshold of prev to define dominant taxa in NCT
lst_NCT <- rprojroot::find_rstudio_root_file() %>% 
  file.path("./data/Chap1/NCTs_v7_wrench.rds") %>% 
  readRDS() %>% 
  get_most_prev(prev_threshold = thres_prev_NCT, N = 1000)
df_additionalInfo <- rprojroot::find_rstudio_root_file() %>%
  file.path("./meta/Mice_meta.xlsx") %>% 
  readxl::read_xlsx()
##---------------------------
pseq_Raw <- rprojroot::find_rstudio_root_file() %>% 
  file.path("data/Chap3", "pseq_Proj5_postFilter_v04.rds") %>% 
  readRDS() %>% 
  ps_filter(ffpe.bulk== "bulk") %>% 
  ps_filter(true.control == "true")
pseq_Decontam <- rprojroot::find_rstudio_root_file() %>% 
  file.path("data/Chap3", "pseq_bulk_decontam_p0.5.rds") %>% 
  readRDS()
pseq_Nj <- rprojroot::find_rstudio_root_file() %>% 
  file.path("./data/Chap3/pseq_bulk_Fisher_v02.rds") %>% 
  readRDS()
pseq_Restrictive <- rprojroot::find_rstudio_root_file() %>% 
  file.path("./data/Chap3/pseq_bulk_restrictive.rds") %>% 
  readRDS()
pseq_SCRuB <- rprojroot::find_rstudio_root_file() %>% 
  file.path("./data/Chap3/pseq_bulk_SCRuB.rds") %>% 
  readRDS()
## change from raw to ND
psLst <- list(ND = pseq_Raw, Restrictive = pseq_Restrictive, Decontam = pseq_Decontam,
              SCRuB = pseq_SCRuB, Nj = pseq_Nj)
##---

```

```{r}
my_order <- c("Nj", "Restrictive", "SCRuB", "Decontam", "ND")
lst_comparison <- list(c("Nj", "Restrictive"), c("Nj", "SCRuB"), c("Nj", "Decontam"), c("Nj", "ND"))
thres_abund <- 0.01
thres_prev <- 0.5
m_by <- "prev"
df_all <- process_survey(psLst = psLst, lstNCT = lst_NCT, m_by = m_by, thres_abund = thres_abund, thres_prev = thres_prev)
df_all$Method <- factor(df_all$Method, levels = my_order)


```

------------------------------------------------------------------------

------------------------------------------------------------------------

**paired Wilcox signed test**

```{r}
plt <- AlphaPlot_Violin_Wilcox_2(df = df_all, SampleID = "SampleID", strata = "Method", val = "Composite_Score", y_label = "Composite Score (Yield x Purity)", title = NULL, subtitle = NULL, my_alternative = "greater")
plt <- plt +
  theme_classic(base_family = "Arial") +
  theme(
    axis.title = element_text(
      family = "Arial",
      size = 12
    ),
    axis.text = element_text(
      family = "Arial",
      size = 12
    ),
    legend.position = "none"
  )
## save to png
ggsave("./Figures_for_Manuscript_files/saved_png/3.B.svg", plot = plt, width = 8, height = 8, dpi = 300, limitsize = F)
plt
```

**Figure 3.B:** Atchiton distance between replicates, statistical test with paired Wilcox ranked test.

------------------------------------------------------------------------

Pairwise p-value

```{r}
mx <- pairwise_WT_Method(df_all ,alternative = "greater", metric = "Composite_Score", strata = "Method", m_order = c("Nj", "Decontam", "Restrictive", "SCRuB", "ND"))
write.csv(mx, "./Figures_for_Manuscript_files/saved_png/Tab_fig3.B_(pairedWilcox).csv", row.names = TRUE)
print(mx)
```

### 3.2 Assessment Using Technical Replicates

Supporting function

```{r}
process_replica <- function(
    ps1, ## original sample
    ps2, ## replica 
    normalize = "compositional", ## compositional - bray | clr - euclidean
    distance = "bray", # jaccard, euclidean
    col_name = "raw"){
  ## add replica info and change names
  ps1 <- ps1 %>% 
    ps_mutate(Replica = "R1") %>% 
    ps_mutate(NewName = sprintf("%s_%s", AN_NR, Replica))
  sample_names(ps1) <- meta(ps1)[["NewName"]]
  
  ps2 <- ps2 %>% 
    ps_mutate(Replica = "R2") %>% 
    ps_mutate(NewName = sprintf("%s_%s", AN_NR, Replica))
  sample_names(ps2) <- meta(ps2)[["NewName"]]
  
  ps_merged <- merge_phyloseq(ps1, ps2)
  if(normalize == "rarefaction"){
    ps_merged <- ps_merged %>% rarefy_even_depth()
  }else if(normalize == "Wrench"){
    ps_merged <- ps_merged %>% WrenchWrapper(grp = "Sex", roundUp = FALSE)
  } else if(normalize %in%  c("clr", "compositional")){
    ps_merged <- ps_merged %>% microbiome::transform(transform = normalize)
  }
  
  # merge_phyloseq(ps1, ps2) %>% 
  #   microbiome::transform(transform = normalize) %>% 
  dis_mx <- NULL
  if(distance == "jaccard"){
    dis_mx <- ps_merged %>% 
      phyloseq::distance(method = distance, binary = TRUE)
  } else{
    dis_mx <- ps_merged %>% 
      phyloseq::distance(method = distance)
  }
  dis_mx <- dis_mx %>% 
    # ps_merged %>% 
    # phyloseq::distance(method = distance) %>% 
    as.matrix() %>% 
    as.data.frame() %>% 
    dplyr::select(contains("R2")) %>% 
    rownames_to_column(var = "rowname") %>% 
    filter(grepl("R1", rowname)) %>% 
    arrange(rowname) %>% 
    column_to_rownames("rowname") %>% 
    dplyr::select(sort(names(.)))
  col <- tibble(AN_NR = sapply(rownames(dis_mx), FUN = function(x){strsplit(x, split = "_")[[1]][1]}), !!sym(col_name) := diag(as.matrix(dis_mx)))
  return(col)
}
##------------------------
process_psLst_with_replica <- function(psLst, dist_metric = "euclidean", trans = "clr"){
  df <- process_replica(ps1 = psLst[["ND_R1"]], ps2 = psLst[["ND_R2"]], normalize = trans, distance = dist_metric, col_name = "ND") %>% 
    left_join(., process_replica(ps1 = psLst[["Restrictive_R1"]], ps2 = psLst[["Restrictive_R2"]], normalize = trans, distance = dist_metric, col_name = "Restrictive"), by = "AN_NR") %>% 
    left_join(., process_replica(ps1 = psLst[["Decontam_R1"]], ps2 = psLst[["Decontam_R2"]], normalize = trans, distance = dist_metric, col_name = "Decontam"), by = "AN_NR") %>% 
    left_join(., process_replica(ps1 = psLst[["SCRuB_R1"]], ps2 = psLst[["SCRuB_R2"]], normalize = trans, distance = dist_metric, col_name = "SCRuB"), by = "AN_NR") %>% 
    left_join(., process_replica(ps1 = psLst[["Nj_R1"]], ps2 = psLst[["Nj_R2"]], normalize = trans, distance = dist_metric, col_name = "Nj"), by = "AN_NR") %>% 
    pivot_longer(cols = -AN_NR, names_to = "Methods", values_to = "IntraDistance")
  return(df)
}
```

Load data and replicates

```{r}
rm(list = setdiff(ls(), c(lsf.str(), "Kolors", "Kolors_Chap2", "df_additionalInfo", "lst_comparison", "my_order")))
## load first replica
df_additionalInfo <- rprojroot::find_rstudio_root_file() %>% 
  file.path("meta/Mice_meta.xlsx") %>% 
  readxl::read_xlsx()
pseq_raw <- rprojroot::find_rstudio_root_file() %>% 
  file.path("data/Chap3/pseq_Proj5_postFilter_v04.rds") %>% 
  readRDS() %>% 
  ps_filter(ffpe.bulk == "bulk") %>% 
  ps_filter(true.control == "true") %>% 
  append_AN_NR(df_additional = df_additionalInfo) #%>% 
  #WrenchWrapper(grp = "Sex") 
pseq_restrictive <- rprojroot::find_rstudio_root_file() %>% 
  file.path("data/Chap3/pseq_bulk_restrictive.rds") %>% 
  readRDS() %>% 
  append_AN_NR(df_additionalInfo) #%>% 
  #WrenchWrapper(grp = "Sex")
pseq_decontam <- rprojroot::find_rstudio_root_file() %>% 
  file.path("data/Chap3/pseq_bulk_decontam_p0.5.rds") %>% 
  readRDS() %>% 
  append_AN_NR(df_additionalInfo) #%>% 
  #WrenchWrapper(grp = "Sex")
pseq_SCRuB <- rprojroot::find_rstudio_root_file() %>%
  file.path("data/Chap3/pseq_bulk_SCRuB.rds") %>% 
  readRDS() %>% 
  append_AN_NR(df_additionalInfo) #%>% 
  #WrenchWrapper(grp = "Sex")
pseq_Nj <- rprojroot::find_rstudio_root_file() %>%
  file.path("data/Chap3/pseq_bulk_Fisher_v02.rds") %>% 
  readRDS() %>% 
  append_AN_NR(df_additionalInfo) #%>% 
  #WrenchWrapper(grp = "Sex")

## load second replica
rep_raw <- rprojroot::find_rstudio_root_file() %>%
  file.path("data/Chap3_Addition/pseq_origin_v1.0.rds") %>% 
  readRDS() %>% 
  ps_filter(ffpe.bulk == "bulk") %>% 
  append_AN_NR(df_additional = df_additionalInfo)
rep_restrictive <- rprojroot::find_rstudio_root_file() %>%
  file.path("data/Chap3_Addition/pseq_restrictive.rds") %>% 
  readRDS() %>% 
  append_AN_NR(df_additional = df_additionalInfo)
rep_decontam <- rprojroot::find_rstudio_root_file() %>%
  file.path("data/Chap3_Addition/pseq_decontam.rds") %>% 
  readRDS() %>% 
  append_AN_NR(df_additional = df_additionalInfo)
rep_SCRuB <- rprojroot::find_rstudio_root_file() %>%
  file.path("data/Chap3_Addition/pseq_SCRuB.rds") %>% 
  readRDS() %>% 
  append_AN_NR(df_additional = df_additionalInfo)
rep_Nj <- rprojroot::find_rstudio_root_file() %>% 
  file.path("data/Chap3_Addition/pseq_Nejman.rds") %>% 
  readRDS() %>% 
  append_AN_NR(df_additional = df_additionalInfo)
## rename taxa
taxa_names(rep_raw) <- gsub(pattern = " ", replacement = "_", x = taxa_names(rep_raw))
taxa_names(rep_restrictive) <- gsub(pattern = " ", replacement = "_", x = taxa_names(rep_restrictive))
taxa_names(rep_decontam) <- gsub(pattern = " ", replacement = "_", x = taxa_names(rep_decontam))
taxa_names(rep_SCRuB) <- gsub(pattern = " ", replacement = "_", x = taxa_names(rep_SCRuB))
taxa_names(rep_Nj) <- gsub(pattern = " ", replacement = "_", x = taxa_names(rep_Nj))

## retain pair
IDs <- meta(rep_raw)[["AN_NR"]]
pseq_raw <- pseq_raw %>% ps_filter(AN_NR %in% IDs)
pseq_restrictive <- pseq_restrictive %>% ps_filter(AN_NR %in% IDs)
pseq_decontam <- pseq_decontam %>% ps_filter(AN_NR %in% IDs)
pseq_SCRuB <- pseq_SCRuB %>% ps_filter(AN_NR %in% IDs)
pseq_Nj <- pseq_Nj %>% ps_filter(AN_NR %in% IDs)
rm(IDs)

## create a list
psLst_2R <- list(ND_R1 = pseq_raw, ND_R2 = rep_raw,
                 Restrictive_R1 = pseq_restrictive, Restrictive_R2 = rep_restrictive,
                 Decontam_R1 = pseq_decontam, Decontam_R2 = rep_decontam,
                 SCRuB_R1 = pseq_SCRuB, SCRuB_R2 = rep_SCRuB,
                 Nj_R1 = pseq_Nj, Nj_R2 = rep_Nj)
rm(list = setdiff(ls(), c(lsf.str(), "Kolors", "Kolors_Chap2", "psLst_2R", "lst_comparison", "my_order")))
## re-arrange sample order
for (i in 1:length(psLst_2R)) {
  psLst_2R[[i]] <- psLst_2R[[i]] %>% ps_arrange(AN_NR)
}
```

#### **Aitchison** - CLR transform with eucledian distance

```{r}
dist_metric <- "euclidean"
trans <-"clr"  #"rarefaction"
df_long <- process_psLst_with_replica(psLst = psLst_2R, dist_metric = dist_metric, trans = trans)
```

Paired Wilcox test

```{r}
#df_long$Methods <- factor(df_long$Methods, levels = c("Nj", "Restrictive", "SCRuB", "Decontam", "ND"))
df_long$Methods <- factor(df_long$Methods, levels = my_order)
plt <- AlphaPlot_Violin_Wilcox_2(df = df_long, SampleID = "AN_NR", strata = "Methods", val = "IntraDistance", y_label = "Inter-replicate Distance", title = NULL, subtitle = NULL, my_alternative = "less") +
  theme_classic(base_family = "Arial") +
  theme(
    axis.title = element_text(
      family = "Arial",
      size = 12
    ),
    axis.text = element_text(
      family = "Arial",
      size = 12
    ),
    legend.position = "none"
  )
##save to png
ggsave("./Figures_for_Manuscript_files/saved_png/3.C.svg", plot = plt, width = 8, height = 6, dpi = 300, limitsize = F)
plt
```

**Figure 3.3**: Inter-replicates distance among decontamination methods.

P-value pairwise

```{r}
mx <- pairwise_WT_Method(df_longer = df_long, metric = "IntraDistance", strata = "Methods", m_alternative = "less", m_order = my_order)
write.csv(mx, "./Figures_for_Manuscript_files/saved_png/Tab_fig3.C_(pairedWilcox).csv", row.names = TRUE)
```

### Jarcard distance with/without rarefaction normalization

```{r}
dist_metric <- "jaccard"
trans <-"identity"  #"rarefaction", Wrench, identity
#dist_metric <- "euclidean"
#trans <-"clr"  #"rarefaction"
df_long <- process_psLst_with_replica(psLst = psLst_2R, dist_metric = dist_metric, trans = trans)
```

Paired Wilcox test

```{r}
df_long$Methods <- factor(df_long$Methods, levels = my_order)
plt <- AlphaPlot_Violin_Wilcox(df = df_long, SampleID = "AN_NR", strata = "Methods", val = "IntraDistance", y_label = "Inter-replicate distance", title = "Paired Wilcox test - unweighted Jaccard", subtitle = "Lower is better", comparisons = lst_comparison, alternative = "less")+
   theme_classic(base_family = "Arial") +
  theme(
    axis.title = element_text(
      family = "Arial",
      size = 11
    ),
    axis.text = element_text(
      family = "Arial",
      size = 8
    ),
    legend.position = "none"
  )
ggsave("./Figures_for_Manuscript_files/saved_png/3_Supplementary_InterReplicate_Jaccard.png", plot = plt, width = 8, height = 6, dpi = 300, limitsize = F)
plt
```

### **3.3 Overlap-Overall Ratio between Replicates**

Supporting functions

```{r}
Samplewise_OverlapOverall_Ratio <- function(psLst){
  
  ## extract taxa name
  for(i in 1:length(psLst)){
    psLst[[i]] <- psLst[[i]] %>%
    ps_mutate(
      taxa_present = purrr::map(
        sample_names(.),
        function(s) {
          otu <- otu_table(.)
          
          # handle taxa orientation
          if (taxa_are_rows(.)) {
            taxa_names(.)[otu[, s] > 0]
          } else {
            taxa_names(.)[otu[s, ] > 0]
          }
        }
      )
    )
  }
  df_Lst <- list()
  for(i in 1:length(psLst)){
    nm <- names(psLst[i])
    df_Lst[[nm]] <- psLst[[i]] %>% 
      meta() %>% 
      remove_rownames() %>% 
      dplyr::select(AN_NR, !!nm := taxa_present) #%>% 
      #mutate(n_taxa = map_int(taxa_present, length)) %>% 
      #rename(!!nm := taxa_present, !!sprintf("n_%s", nm) := n_taxa)
  }
  ##
  ## merge metadata
  df_taxa <- Reduce(function(x,y) merge(x,y, by = "AN_NR"), df_Lst)
  Pair_names <- list(c("ND_R1","ND_R2"), c("Restrictive_R1","Restrictive_R2"),
                     c("Decontam_R1", "Decontam_R2"), c("SCRuB_R1", "SCRuB_R2"),
                     c("Nj_R1", "Nj_R2"))
  for (pair in Pair_names) {
    col1 <- pair[1]
    col2 <- pair[2]
    nm <- strsplit(col1, split = "_")[[1]][1]
    intersect_col <- sprintf("%s_intersect", nm)
    union_col <- sprintf("%s_union", nm)
    df_taxa <- df_taxa %>% 
      mutate(
        !!intersect_col := map2(!!sym(col1), !!sym(col2), intersect),
        !!union_col := map2(!!sym(col1), !!sym(col2), union)
      )
  }
  ## list-column
  list_cols <- names(df_taxa)[map_lgl(df_taxa, is.list)]
  for(col in list_cols){
    new_col <- paste0("n_", col)
    df_taxa <- df_taxa %>% 
      mutate(!!new_col := map_int(!!sym(col), length))
  }
  #####------------
  df_num <- df_taxa %>% 
    # mutate(Raw_intersect = map2(Raw_R1, Raw_R2, intersect),
    #        Raw_union = map2(Raw_R1, Raw_R2, union),
    #        n_Raw_intersect = map_int(Raw_intersect, length),
    #        n_Raw_union = map_int(Raw_union, length)) %>% 
    dplyr::select(AN_NR, starts_with("n_")) %>% 
    mutate(
      ND = n_ND_intersect/n_ND_union,
      Restrictive = n_Restrictive_intersect/n_Restrictive_union,
      Decontam = n_Decontam_intersect/n_Decontam_union,
      SCRuB = n_SCRuB_intersect/n_SCRuB_union,
      Nj = n_Nj_intersect/n_Nj_union
         ) %>% 
    dplyr::select(AN_NR, ND, Restrictive, Decontam, SCRuB, Nj)
  return(df_num)
}
### function to filter taxa
Reduce_phyloseq <- function(psLst, rank_level = "species", min_prev = 0.1, min_abund = 0.0001){
  ## check if need tax_agg
  if(rank_level !="species"){
    for(i in 1:length(psLst)){
      psLst[[i]] <- psLst[[i]] %>% 
        tax_agg(rank = rank_level)
    }
  }
  for(i in 1:length(psLst)){
    psLst[[i]] <- psLst[[i]] %>% 
      ps_get() %>% 
      microViz::tax_filter(min_prevalence = min_prev, min_total_abundance = min_abund)
  }
  return(psLst)
}
```

Wilcox test

```{r}
df_long <- Samplewise_OverlapOverall_Ratio(psLst = psLst_2R) %>% 
  pivot_longer(cols = -AN_NR, names_to = "Method", values_to = "Ratio") 
df_long$Method <- factor(df_long$Method, levels = my_order)
plt <- df_long %>% 
  AlphaPlot_Violin_Wilcox_2(SampleID = "AN_NR", strata = "Method", val = "Ratio", y_label = "Compositional Overlap Ratio", title = NULL, subtitle = NULL, my_alternative = "greater") +
  theme_classic(base_family = "Arial") +
  theme(
    axis.title = element_text(
      family = "Arial",
      size = 12
    ),
    axis.text = element_text(
      family = "Arial",
      size = 12
    ),
    legend.position = "none"
  )
## save to png
ggsave("./Figures_for_Manuscript_files/saved_png/3.D.svg", plot = plt, width = 8, height = 8, dpi = 300, limitsize = F)
plt
```

**Figure 3.4**: Proportion between overlap and overall number of taxa between replicates for each methods.

Pairwise table

```{r}
mx <- pairwise_WT_Method(df_longer = df_long, metric = "Ratio", strata = "Method", m_alternative = "greater", m_order = my_order)
write.csv(mx, "./Figures_for_Manuscript_files/saved_png/Tab_fig3.D_Ratio.csv", row.names = TRUE)
```

------------------------------------------------------------------------

##### Filter rare taxa in another rank

```{r}
min_prev <- 0.2
min_abund <- 0.0001
rnk_level <- "genus"
df_long <- Reduce_phyloseq(psLst_2R, rank_level = rnk_level, min_abund = min_abund, min_prev = min_prev) %>% 
  Samplewise_OverlapOverall_Ratio() %>% 
  pivot_longer(cols = -AN_NR, names_to = "Method", values_to = "Ratio") 
df_long$Method <- factor(df_long$Method, levels = c("Nj", "Restrictive", "SCRuB", "Decontam", "Raw"))
plt <- df_long %>% 
  AlphaPlot_Violin_Wilcox(SampleID = "AN_NR", strata = "Method", val = "Ratio", y_label = "Overlap-Oerall Ratio", title = "Paired Wilcox test of Ratio", subtitle = "Higher is better", comparisons = list(c("Nj", "Restrictive"), c("Nj", "SCRuB"), c("Nj", "Decontam"), c("Nj", "Raw")), alternative = "greater")
ggsave("./Figures_for_Manuscript_files/saved_png/3.D.png", plot = plt, width = 8, height = 6, dpi = 300)

## save pairwise into a table
mx <- pairwise_WT_Method(df_longer = df_long, metric = "Ratio", strata = "Method", alternative = "greater", m_order = c("Nj", "Restrictive", "SCRuB", "Decontam", "Raw"))
write.csv(mx, "./Figures_for_Manuscript_files/saved_png/Tab_fig3.D_Ratio.csv", row.names = TRUE)
```

```{r}
min_prev <- 0.1
min_abund <- 0.0001
rnk_level <- "genus"
df_long <- Reduce_phyloseq(psLst_2R, rank_level = rnk_level, min_abund = min_abund, min_prev = min_prev) %>% 
  Samplewise_OverlapOverall_Ratio() %>% 
  pivot_longer(cols = -AN_NR, names_to = "Method", values_to = "Ratio") 
df_long$Method <- factor(df_long$Method, levels = c("Nj", "Restrictive", "SCRuB", "Decontam", "Raw"))
df_long %>% 
  AlphaPlot_Violin_Wilcox(SampleID = "AN_NR", strata = "Method", val = "Ratio", y_label = "Overlap-Oerall Ratio", title = "Paired Wilcox test of Ratio", subtitle = "Higher is better", comparisons = list(c("Nj", "Restrictive"), c("Nj", "SCRuB"), c("Nj", "Decontam"), c("Nj", "Raw")), alternative = "greater")
```

## Chapter 3E

PDAC Intratumoral microbiome investigation/confirmation

Supporting functions

```{r}
rename_samples <- function(ps, m_Replica = "R1", m_Method = "Raw", trans = "identity"){
  ps <- ps %>% 
    ps_mutate(Replica = m_Replica, Method = m_Method) %>% 
    ps_mutate(Group = sprintf("%s_%s", Method, Replica)) %>% 
    ps_mutate(NewID = sprintf("%s_%s", Group, AN_NR)) %>% ps_arrange(AN_NR) %>% 
    microbiome::transform(transform = trans)
  sample_names(ps) <- meta(ps)[["NewID"]]
  return(ps)
}
## process overlap
overlap_ps <- function(ps1, ps2, m_group = "AN_NR"){
  shared_taxa <- intersect(taxa_names(ps1), taxa_names(ps2))
    ol_ps <- merge_phyloseq(ps1 %>% prune_taxa(taxa_names(.) %in% shared_taxa, .) %>% microbiome::transform(transform = "compositional"),
                           ps2 %>% prune_taxa(taxa_names(.) %in% shared_taxa, .) %>% microbiome::transform(transform = "compositional")
                          ) %>% 
    merge_samples(group = m_group)
  AN_lst <- sample_names(ol_ps)
  
  ol_ps <- ol_ps %>% 
    ps_mutate(Replica = "Overlap", Method = meta(ps1)[["Method"]][[1]], AN_NR = AN_lst) %>% 
    ps_mutate(Group = sprintf("%s_%s", Method, Replica)) %>% 
    ps_mutate(NewID = sprintf("%s_%s", Group, AN_NR)) %>% ps_arrange(AN_NR)
  sample_names(ol_ps) <- meta(ol_ps)[["NewID"]]
  return(ol_ps)
}
```

Clean data

```{r}
rm(list = setdiff(ls(), c(lsf.str(), "Kolors", "Kolors_Chap2", "psLst_2R", "my_order", "lst_comparison")))
for(i in 1:length(psLst_2R)){
  tmp <- names(psLst_2R[i]) %>% strsplit(split = "_") %>% first() 
  psLst_2R[[i]] <- rename_samples(psLst_2R[[i]], m_Replica = tmp[2], m_Method = tmp[1])
}
ol_ND <- overlap_ps(psLst_2R$ND_R1, psLst_2R$ND_R2, m_group = "AN_NR")
ol_restrictive <- overlap_ps(psLst_2R$Restrictive_R1, psLst_2R$Restrictive_R2, m_group = "AN_NR")
ol_decontam <- overlap_ps(psLst_2R$Decontam_R1, psLst_2R$Decontam_R2, m_group = "AN_NR")
ol_SCRuB <- overlap_ps(psLst_2R$SCRuB_R1, psLst_2R$SCRuB_R2, m_group = "AN_NR")
ol_Nj <- overlap_ps(psLst_2R$Nj_R1, psLst_2R$Nj_R2, m_group = "AN_NR")

ps_all <- merge_phyloseq(psLst_2R$ND_R1, psLst_2R$ND_R2,
                         psLst_2R$Restrictive_R1, psLst_2R$Restrictive_R2,
                         psLst_2R$Decontam_R1, psLst_2R$Decontam_R2,
                         psLst_2R$SCRuB_R1, psLst_2R$SCRuB_R2,
                         psLst_2R$Nj_R1, psLst_2R$Nj_R2,
                         ol_ND, ol_restrictive, ol_decontam, ol_SCRuB, ol_Nj)
df_meta <- meta(ps_all) %>% 
  group_by(Group) %>% 
  summarise(
    across(
      everything(),
      ~ if (n_distinct(.) == 1) first(.) else NA,
      .names = "{.col}"
    ),
    .groups = "drop"
  ) %>% 
  mutate(NewID = Group, AN_NR = Replica) %>% 
  remove_rownames() %>% 
  column_to_rownames(var = "NewID")
ps_2 <- ps_all %>% 
  merge_samples(group = "Group") ## %>% 
sample_data(ps_2) <- df_meta
```

Rank: genus

```{r}
sample_data(ps_2)$Method <- factor(sample_data(ps_2)$Method, levels = my_order) 
plt <- ps_2 %>% 
   comp_barplot(
    tax_level = "genus", n_taxa = 15, bar_width = 0.7, sample_order = "asis", # don't sort
    x = "Replica" # x argument is available since microViz 0.9.7
  ) +
  facet_wrap(
    facets = vars(Method),
    scales = "fixed"
  ) +
  coord_flip() +
  labs(x = "Sample ID", y = "Relative abundance") +
  scale_y_continuous(expand = expansion(add = c(0, 0.05))) + # axis starts exactly at 0
  theme_bw() + # slightly clearer axes for facets
  theme(panel.spacing.x = unit(6, "mm")) # space for non-overlapping axis numbers
ggsave("Figures_for_Manuscript_files/saved_png/3.E.svg", plot = plt, width = 8, height = 6, dpi = 300)
plt
```

**Figure 4.1**: Bacterial composition of two replicates and their overlap, on original data and with various decontamination methods.

## Chapter 4

Intratumoral microbiome study for low-biomass PDAC

Load FFPE data

```{r}
df_additionalInfo <- rprojroot::find_rstudio_root_file() %>% 
  file.path("meta/Mice_meta.xlsx") %>% 
  readxl::read_xlsx()
IDs <- meta(psLst_2R$ND_R1)[["AN_NR"]]
ffpe_ND <- rprojroot::find_rstudio_root_file() %>% 
  file.path("data/Chap3/pseq_Proj5_postFilter_v04.rds") %>% 
  readRDS() %>% 
  ps_filter(ffpe.bulk == "FFPE") %>% 
  ps_filter(true.control == "true") %>% 
  ps_filter(AN_NR %in% IDs) %>% 
  append_AN_NR(df_additional = df_additionalInfo) %>% 
  ps_arrange(AN_NR)
ffpe_restrictive <- rprojroot::find_rstudio_root_file() %>% 
  file.path("data/Chap3/pseq_FFPE_restrictive.rds") %>% 
  readRDS() %>% 
  ps_filter(AN_NR %in% IDs) %>% 
  append_AN_NR(df_additionalInfo) %>% 
  ps_arrange(AN_NR)
ffpe_decontam <- rprojroot::find_rstudio_root_file() %>% 
  file.path("data/Chap3/pseq_FFPE_decontam.rds") %>% 
  readRDS() %>% 
  ps_filter(AN_NR %in% IDs) %>% 
  append_AN_NR(df_additionalInfo) %>% 
  ps_arrange(AN_NR)
ffpe_SCRuB <- rprojroot::find_rstudio_root_file() %>% 
  file.path("data/Chap3/pseq_FFPE_SCRuB.rds") %>% 
  readRDS() %>% 
  ps_filter(AN_NR %in% IDs) %>% 
  append_AN_NR(df_additionalInfo) %>% 
  ps_arrange(AN_NR)
ffpe_Nj <- rprojroot::find_rstudio_root_file() %>% 
  file.path("data/Chap3/pseq_FFPE_Biom.rds") %>% 
  readRDS() %>% 
  ps_filter(AN_NR %in% IDs) %>% 
  append_AN_NR(df_additionalInfo) %>% 
  ps_arrange(AN_NR)
psLst_3R <- c(psLst_2R, list(ND_ffpe = ffpe_ND, Restrictive_ffpe = ffpe_restrictive,
                             Decontam_ffpe = ffpe_decontam, SCRuB_ffpe = ffpe_SCRuB,
                             Nj_ffpe = ffpe_Nj))
rm(list = setdiff(ls(), c(lsf.str(), "Kolors", "Kolors_Chap2", "psLst_3R", "my_order", "lst_comparison")))
```

**Plots for Alpha and Beta Analyses**

```{r}
for(i in 1:length(psLst_3R)){
  tmp <- names(psLst_3R[i]) %>% strsplit(split = "_") %>% first() 
  psLst_3R[[i]] <- rename_samples(psLst_3R[[i]], m_Replica = tmp[2], m_Method = tmp[1])
}
ps <- merge_phyloseq(psLst_3R[[1]], psLst_3R[[2]], psLst_3R[[3]],psLst_3R[[4]],psLst_3R[[5]],
                     psLst_3R[[6]], psLst_3R[[7]], psLst_3R[[8]],psLst_3R[[9]],psLst_3R[[10]],
                     psLst_3R[[11]], psLst_3R[[12]], psLst_3R[[13]],psLst_3R[[14]],psLst_3R[[15]])
combined <- expand_grid(Index = c("Observed", "Shannon"), Method = c("ND", "Restrictive", "Decontam", "SCRuB", "Nj"))
arr_plots <- apply(combined, 1, FUN = function(x){
  y_lab = ""
  x_lab = ""
  if(x[["Method"]] == "ND"){
    if(x[["Index"]] == "Observed"){
      y_lab = "Number of species"
    }else{
      y_lab = "Shannon index"
    }
  }
  ps %>% 
    ps_filter(Method == x[2]) %>% 
    AlphaPlot_Violin(index = x[1], strata = "Replica", y_label = y_lab, add_legend = T, x_label = "", m_Kolors = Kolors)
})

# Arrange into rows
row1 <- wrap_plots(arr_plots[1:5], nrow = 1)
row2 <- wrap_plots(arr_plots[6:10], nrow = 1)

# Column labels
col_labels <- wrap_plots(
  lapply(c("ND", "Restrictive", "Decontam", "SCRuB", "Nj"), function(x)
    wrap_elements(full = grid::textGrob(
      x,
      gp = grid::gpar(fontsize = 14, fontface = "bold")
    ))
  ),
  nrow = 1
)

# Row labels
row1_label <- wrap_elements(full = grid::textGrob(
  "Observed",
  rot = 90,
  gp = grid::gpar(fontsize = 14, fontface = "bold")
))

row2_label <- wrap_elements(full = grid::textGrob(
  "Shannon",
  rot = 90,
  gp = grid::gpar(fontsize = 14, fontface = "bold")
))

# Final layout
final_plot <-
  col_labels /
  ((row1_label | row1) /
   (row2_label | row2)) +
  plot_layout(
    guides = "collect"
    # heights = c(0.08, 1),
    # widths = c(0.06, 1)
  )

ggsave("Figures_for_Manuscript_files/saved_png/4.C.svg",
       final_plot,
       width = 5 * 3.2,
       height = 2 * 3.6,
       limitsize = FALSE, dpi = 300)

# plt1 <- ps %>% 
#   ps_filter(Method == "ND") %>% 
#   AlphaPlot_Violin(index = "Observed", strata = "Replica", y_label = "Number of species", add_legend = T, gg_title = "", m_Kolors = Kolors, x_label = "Replica")
# plt2 <- AlphaPlot_Violin(ps, index = "Shannon", strata = "Replica", y_label = "Shannon index", add_legend = T, gg_title = "", m_Kolors = Kolors, x_label = "Replica")
# plt <- plt1 + plt2 + plot_layout(guides = "collect")
# ggsave("./Figures_for_Manuscript_files/saved_png/4.C.svg", plt, width = 8, height = 6, dpi = 300)

## beta
plt_beta <- ps %>% 
  ps_filter(Method == "SCRuB") %>%
  BetaPlot(strata_f = "Replica", n_per = 9999, title_method = "bray-distance", dis_method = "bray", ordination_method = "PCoA")
```

Plot alpha + continue

```{r}
methods <- c("ND", "Restrictive", "Decontam", "SCRuB", "Nj")

corner <- plot_spacer()

top_labels <- lapply(methods, function(x)
  wrap_elements(full = textGrob(x, gp = gpar(fontsize = 14, fontface = "bold")))
)

row1_label <- wrap_elements(full = textGrob("Observed", rot = 90,
                                             gp = gpar(fontsize = 14, fontface = "bold")))
row2_label <- wrap_elements(full = textGrob("Shannon", rot = 90,
                                             gp = gpar(fontsize = 14, fontface = "bold")))

# Flatten everything into ONE list, in row-major order
all_panels <- c(
  list(corner), top_labels,             # row 1: spacer + 5 column labels
  list(row1_label), arr_plots[1:5],     # row 2: label + 5 Observed plots
  list(row2_label), arr_plots[6:10]     # row 3: label + 5 Shannon plots
)

# 6 columns (label col + 5 method cols) x 3 rows
design1 <- "
ABCDEF
GHIJKL
MNOPQR
"

final_plot <- wrap_plots(all_panels) +
  plot_layout(
    design  = design1,
    heights = c(0.08, 1, 1),
    widths  = c(0.06, 1, 1, 1, 1, 1),
    guides  = "collect"
  )

ggsave("Figures_for_Manuscript_files/saved_png/4.C.svg",
       final_plot,
       width = 5 * 3.2,
       height = 2 * 3.6,
       limitsize = FALSE, dpi = 300)
```

Beta plots

```{r}
## beta
# plt_beta <- ps %>% 
#   ps_filter(Method == "SCRuB") %>%
#   BetaPlot(strata_f = "Replica", n_per = 9999, title_method = "bray-distance", dis_method = "bray", ordination_method = "PCoA")
normalize_methods <- c("Wrench", "Rarefaction")
combined <- expand_grid(Normalization = normalize_methods, Method = c("ND", "Restrictive", "Decontam", "SCRuB", "Nj"))

beta_plots <- apply(combined, 1, function(x){
  ps_tmp <- ps %>% 
    ps_filter(Method == x[["Method"]])
  if(x[["Normalization"]] == "Wrench"){
    ps_tmp <- ps_tmp %>% 
      WrenchWrapper(grp = "Replica")
  } else{
    ps_tmp <- ps_tmp %>% 
      rarefy_even_depth()
  }
  ps_tmp %>% 
    BetaPlot(strata_f = "Replica", n_per = 9999, title_method = "", dis_method = "bray", ordination_method = "PCoA")
})

methods <- c("ND", "Restrictive", "Decontam", "SCRuB", "Nj")

corner <- plot_spacer()

top_labels <- lapply(methods, function(x)
  wrap_elements(full = textGrob(x, gp = gpar(fontsize = 14, fontface = "bold")))
)

row1_label <- wrap_elements(full = textGrob("Wrench", rot = 90,
                                             gp = gpar(fontsize = 14, fontface = "bold")))
row2_label <- wrap_elements(full = textGrob("Rarefaction", rot = 90,
                                             gp = gpar(fontsize = 14, fontface = "bold")))

# Flatten everything into ONE list, in row-major order
all_panels <- c(
  list(corner), top_labels,             # row 1: spacer + 5 column labels
  list(row1_label), beta_plots[1:5],     # row 2: label + 5 Observed plots
  list(row2_label), beta_plots[6:10]     # row 3: label + 5 Shannon plots
)

# 6 columns (label col + 5 method cols) x 3 rows
design1 <- "
ABCDEF
GHIJKL
MNOPQR
"

final_plot <- wrap_plots(all_panels) +
  plot_layout(
    design  = design1,
    heights = c(0.08, 1, 1),
    widths  = c(0.06, 1, 1, 1, 1, 1),
    guides  = "collect"
  )
ggsave("Figures_for_Manuscript_files/saved_png/4.D.svg",
       final_plot,
       width = 5 * 3.2,
       height = 2 * 3.6,
       limitsize = FALSE, dpi = 300)
```

Pairwise - table

```{r}
## pairwise : FF rep1, FF rep2, FFPE
normalize_methods <- c("Wrench", "Rarefaction")
combined <- expand_grid(Normalization = normalize_methods, Method = c("ND", "Restrictive", "Decontam", "SCRuB", "Nj"))

beta_adonis2_pairwise <- apply(combined, 1, function(x){
  ps_tmp <- ps %>% 
    ps_filter(Method == x[["Method"]])
  if(x[["Normalization"]] == "Wrench"){
    ps_tmp <- ps_tmp %>% 
      WrenchWrapper(grp = "Replica")
  } else{
    ps_tmp <- ps_tmp %>% 
      rarefy_even_depth()
  }
  ps_tmp %>% 
    BetaPlot_pairwise(grp = "Replica", n_per = 9999, title_method = "", dis_method = "bray", ordination_method = "PCoA")
})

## R2
R_Square <- combined
R_Square$R1_vs_R2 <- lapply(beta_adonis2_pairwise, function(x){x[["R1_vs_R2"]][["R2"]][[1]]}) %>% unlist()
R_Square$R1_vs_FFPE <- lapply(beta_adonis2_pairwise, function(x){x[["R1_vs_ffpe"]][["R2"]][[1]]}) %>% unlist()
R_Square$R2_vs_FFPE <- lapply(beta_adonis2_pairwise, function(x){x[["R2_vs_ffpe"]][["R2"]][[1]]}) %>% unlist()
R_Square <- R_Square %>% 
  pivot_longer(cols = -c("Normalization", "Method"), names_to = "Pairs", values_to = "R2")

## Pr(>F)
P_val <- combined
P_val$R1_vs_R2 <- lapply(beta_adonis2_pairwise, function(x){x[["R1_vs_R2"]][["Pr(>F)"]][[1]]}) %>% unlist()
P_val$R1_vs_FFPE <- lapply(beta_adonis2_pairwise, function(x){x[["R1_vs_ffpe"]][["Pr(>F)"]][[1]]}) %>% unlist()
P_val$R2_vs_FFPE <- lapply(beta_adonis2_pairwise, function(x){x[["R2_vs_ffpe"]][["Pr(>F)"]][[1]]}) %>% unlist()
P_val <- P_val %>% 
  pivot_longer(cols = -c("Normalization", "Method"), names_to = "Pairs", values_to = "p_value")
# combine together
res <- left_join(R_Square, P_val, by = c("Normalization", "Method", "Pairs"))
write_csv(res, "./Figures_for_Manuscript_files/saved_png/Chap4_beta_pairwise.csv")
```

**Illustration**

### Grid of Venn diagram

```{r}
library(ggVennDiagram)
library(patchwork)
library(stringr)
library(ggplot2)

## -----------------------------------------------------------------
## 0. Parse method / replicate from names like "Method1_R1"
## -----------------------------------------------------------------
## rename phyloseq
for(i in 1:length(psLst_3R)){
  sample_names(psLst_3R[[i]]) <- meta(psLst_3R[[i]])$AN_NR
}

meta_df <- data.frame(
  name      = names(psLst_3R),
  method    = str_extract(names(psLst_3R), "^[^_]+"),   # "Method1", "Method2", ...
  replicate = str_extract(names(psLst_3R), "(?<=_).+$"),   # "R1", "R2", "R3"
  stringsAsFactors = FALSE
)

methods_all <- unique(meta_df$method)
stopifnot(length(methods_all) == 5)

## -----------------------------------------------------------------
## 1. Helper: get taxa present (per replicate) for one method + sample
## -----------------------------------------------------------------
get_taxa_sets <- function(method, sample_id, psLst_3R, meta_df) {
  sub <- meta_df[meta_df$method == method, ]
  sub <- sub[order(sub$replicate), ]   # ensures R1, R2, R3 order

  taxa_sets <- lapply(sub$name, function(nm) {
    ps <- psLst_3R[[nm]]
    otu <- as(otu_table(ps), "matrix")
    if (!taxa_are_rows(ps)) otu <- t(otu)

    if (!(sample_id %in% colnames(otu))) {
      stop(paste("Sample", sample_id, "not found in", nm))
    }
    abund <- otu[, sample_id]
    taxa_names(ps)[abund > 0]
  })

  names(taxa_sets) <- sub$replicate
  taxa_sets
}

## -----------------------------------------------------------------
## 2. Sample IDs (assumes all 15 phyloseq objects share same 10 samples)
## -----------------------------------------------------------------
sample_ids <- sample_names(psLst_3R[[1]])
stopifnot(length(sample_ids) == 10)

## -----------------------------------------------------------------
## 3. Venn diagram for one (sample, method) cell
## -----------------------------------------------------------------
make_venn <- function(sample_id, method) {
  taxa_sets <- get_taxa_sets(method, sample_id, psLst_3R, meta_df)

  ggVennDiagram(taxa_sets,
                label = "count",
                edge_size = 0.3) +
    scale_fill_gradient(low = "white", high = "steelblue") +
    theme_void() +
    theme(legend.position = "none",
          plot.margin = margin(1, 1, 1, 1))
}

## -----------------------------------------------------------------
## 4. Build all 50 plots (10 samples x 5 methods), row-major order
## -----------------------------------------------------------------
plot_grid_list <- list()
idx <- 1
for (s in sample_ids) {
  for (m in methods_all) {
    plot_grid_list[[idx]] <- make_venn(s, m)
    idx <- idx + 1
  }
}

## -----------------------------------------------------------------
## 5. Assemble grid with column headers (methods) and row headers (samples)
## -----------------------------------------------------------------
# Column header row
col_titles <- wrap_plots(
  lapply(methods_all, function(m) {
    ggplot() + theme_void() +
      annotate("text", x = 0, y = 0, label = m, fontface = "bold", size = 3.5)
  }),
  ncol = length(methods_all)
)

# Main grid, one row per sample, with sample ID label prepended on the left
build_row <- function(sample_id) {
  label_plot <- ggplot() + theme_void() +
    annotate("text", x = 0, y = 0, label = sample_id, fontface = "bold",
             angle = 90, size = 3.2)

  row_plots <- lapply(methods_all, function(m) plot_grid_list[[
    which(rep(sample_ids, each = length(methods_all)) == sample_id &
          rep(methods_all, times = length(sample_ids)) == m)
  ]])

  wrap_plots(c(list(label_plot), row_plots),
             ncol = length(methods_all) + 1,
             widths = c(0.15, rep(1, length(methods_all))))
}

all_rows <- lapply(sample_ids, build_row)

grid_body <- wrap_plots(all_rows, ncol = 1)

final_plot <- (
  wrap_plots(c(list(ggplot() + theme_void()), 
               lapply(methods_all, function(m) {
                 ggplot() + theme_void() +
                   annotate("text", x = 0, y = 0, label = m, fontface = "bold", size = 3.5)
               })),
             ncol = length(methods_all) + 1,
             widths = c(0.15, rep(1, length(methods_all))))
  / grid_body
) + plot_layout(heights = c(0.04, 1))

## -----------------------------------------------------------------
## 6. Save
## -----------------------------------------------------------------
ggsave("4.B.svg",
       final_plot,
       width = 5 * 3.2,
       height = 10 * 2.4,
       limitsize = FALSE)
```

### Barplot: FF1, FF2, OverlapFF, FFPE

prepare

```{r}
for(i in 1:length(psLst_3R)){
  tmp <- names(psLst_3R[i]) %>% strsplit(split = "_") %>% first() 
  psLst_3R[[i]] <- rename_samples(psLst_3R[[i]], m_Replica = tmp[2], m_Method = tmp[1], trans = "identity")
}
ol_ND <- overlap_ps(psLst_3R$ND_R1, psLst_3R$ND_R2, m_group = "AN_NR")
ol_restrictive <- overlap_ps(psLst_3R$Restrictive_R1, psLst_3R$Restrictive_R2, m_group = "AN_NR")
ol_decontam <- overlap_ps(psLst_3R$Decontam_R1, psLst_3R$Decontam_R2, m_group = "AN_NR")
ol_SCRuB <- overlap_ps(psLst_3R$SCRuB_R1, psLst_3R$SCRuB_R2, m_group = "AN_NR")
ol_Nj <- overlap_ps(psLst_3R$Nj_R1, psLst_3R$Nj_R2, m_group = "AN_NR")

ps_all <- merge_phyloseq(psLst_3R$ND_R1 %>% microbiome::transform(transform = "compositional"), 
                         psLst_3R$ND_R2 %>% microbiome::transform(transform = "compositional"), 
                         psLst_3R$ND_ffpe %>% microbiome::transform(transform = "compositional"),
                         psLst_3R$Restrictive_R1 %>% microbiome::transform(transform = "compositional"), 
                         psLst_3R$Restrictive_R2 %>% microbiome::transform(transform = "compositional"), 
                         psLst_3R$Restrictive_ffpe %>% microbiome::transform(transform = "compositional"),
                         psLst_3R$Decontam_R1 %>% microbiome::transform(transform = "compositional"), 
                         psLst_3R$Decontam_R2 %>% microbiome::transform(transform = "compositional"), 
                         psLst_3R$Decontam_ffpe %>% microbiome::transform(transform = "compositional"),
                         psLst_3R$SCRuB_R1 %>% microbiome::transform(transform = "compositional"), 
                         psLst_3R$SCRuB_R2 %>% microbiome::transform(transform = "compositional"), 
                         psLst_3R$SCRuB_ffpe %>% microbiome::transform(transform = "compositional"),
                         psLst_3R$Nj_R1 %>% microbiome::transform(transform = "compositional"), 
                         psLst_3R$Nj_R2 %>% microbiome::transform(transform = "compositional"), 
                         psLst_3R$Nj_ffpe %>% microbiome::transform(transform = "compositional"),
                         ol_ND %>% microbiome::transform(transform = "compositional"), 
                         ol_restrictive %>% microbiome::transform(transform = "compositional"), 
                         ol_decontam %>% microbiome::transform(transform = "compositional"), 
                         ol_SCRuB %>% microbiome::transform(transform = "compositional"), 
                         ol_Nj %>% microbiome::transform(transform = "compositional"))
df_meta <- meta(ps_all) %>% 
  group_by(Group) %>% 
  summarise(
    across(
      everything(),
      ~ if (n_distinct(.) == 1) first(.) else NA,
      .names = "{.col}"
    ),
    .groups = "drop"
  ) %>% 
  mutate(NewID = Group, AN_NR = Replica) %>% 
  remove_rownames() %>% 
  column_to_rownames(var = "NewID")
ps_2 <- ps_all %>% 
  merge_samples(group = "Group") ## %>% 
sample_data(ps_2) <- df_meta
```

plots

```{r}
sample_data(ps_2)$Method <- factor(sample_data(ps_2)$Method, levels = c("ND", "Restrictive", "Decontam", "SCRuB", "Nj"))
plt <- ps_2 %>% 
  ps_filter(Replica %in% c("ffpe", "Overlap")) %>% ## only keep ND and Nj
   comp_barplot(
    tax_level = "species", n_taxa = 20, bar_width = 0.7, sample_order = "asis", # don't sort
    x = "Replica" # x argument is available since microViz 0.9.7
  ) +
  facet_wrap(
    facets = vars(Method),
    scales = "fixed",
    nrow = 1
  ) +
  coord_flip() +
  labs(x = "Sample ID", y = "Relative abundance") +
  scale_y_continuous(expand = expansion(add = c(0, 0.05))) + # axis starts exactly at 0
  theme_bw() + # slightly clearer axes for facets
  theme_classic(base_family = "Arial") + 
  theme(panel.spacing.x = unit(6, "mm"), # space for non-overlapping axis numbers
        axis.title = element_text(
      family = "Arial",
      size = 11
    ),
    axis.text = element_text(
      family = "Arial",
      size = 8
    ),
    legend.position = "bottom"   # <- species legend at bottom
  ) 
ggsave("Figures_for_Manuscript_files/saved_png/4.A_FFPE_Overlap_species.svg", plot = plt, width = 9, height = 6, dpi = 300, limitsize = FALSE)
plt
```
