---
title: "Figures for Manuscript"
author: "Linh Dang"
editor: visual
execute:
  warning: false
  message: false
format:
  html:
    code-fold: true
    toc: true
    toc-location: left
    toc-depth: 3
    fig-responsive: true
    fig-lightbox: true
---

## Introduction

This is a script to recreate figures for the manuscript.

**Library**

```{r}
#| message: false

library(phyloseq)
library(reshape2)
library(microViz)
library(microbiome)
library(vegan)
library(lme4)       # Linear Mixed Models (for repeated measures)
library(lmerTest)   # P-values for Mixed Models
library(emmeans)    # Pairwise comparisons
library(Wrench)
library(rstatix)
library(ggpubr)
library(tidyverse)
library(VennDiagram)
library(ComplexUpset)
library(patchwork)
library(pairwiseAdonis)
library(MASS)
rm(list = ls())
```

## Figures for LP-Environment Investigation

------------------------------------------------------------------------

Load data

```{r}
source("./Functions.R")
## remove old object
rm(list = setdiff(ls(), c(lsf.str(), "Kolors", "Kolors_Chap2")))
pseq <- rprojroot::find_rstudio_root_file() %>% 
  file.path("data/Chap2/px_J_U_NCT_v02.rds") %>% 
  readRDS() %>% 
  ps_get() %>% 
  tax_filter(min_total_abundance = 2, min_prevalence = 2)
## edit meta
pseq <- pseq %>% 
  ps_mutate(LP = ifelse(person == "J", "LP1", "LP2")) %>% 
  ps_mutate(Environment = HOOD_BENCH) %>% 
  ps_select(-c("HOOD_BENCH", "person"))
```

### Alpha

```{r}
lst <- AlphaPlotWrapper_Violin(pseq,roundUp = TRUE,  strata = "LP", m_Kolors = Kolors_Chap2, x_label = "LP")
# plt <- lst$Observed + lst$Shannon + lst$InvSimpson + plot_layout(guides = "collect")
# ggsave("./Figures_for_Manuscript_files/saved_png/2.B.svg", plot = plt, width = 8, height = 6, dpi = 300)
# plt

## Alternative
al1 <- AlphaPlot_Violin_glm(pseq, index = "Observed", main_var = "Environment", covariate_var = "LP", x_label = "Environment", y_label = "Number of observed species", add_legend = T, gg_title = NULL, gg_subtitle = NULL, m_Kolors = Kolors_Chap2[3:4], m_family = NULL, is_nb = T)
al2 <- AlphaPlot_Violin_glm(pseq, index = "Shannon", main_var = "Environment", covariate_var = "LP", x_label = "Environment", y_label = "Shannon index", add_legend = T, gg_title = NULL, gg_subtitle = NULL, m_Kolors = Kolors_Chap2[3:4], m_family = gaussian(), is_nb = F)

plt <- al1$plt + al2$plt + plot_layout(guides = "collect")
plt
#ggsave("./Figures_for_Manuscript_files/saved_png/S2.B.svg", plot = plt, width = 8, height = 6, dpi = 300)
  
# al2$const %>% 
#   summary(infer = c(TRUE, TRUE)) %>% 
#   as.data.frame() %>% 
#   mutate(exp = exp(estimate),
#          LCL = exp(lower.CL),
#          UCL = exp(upper.CL)) %>% 
#   dplyr::select(contrast, estimate, exp, LCL, UCL, p.value) %>% 
#   write_csv("tmp.txt")
```

**Figure 2B**: Alpha diversity of NCT samples stratified by persons who processed samples.

------------------------------------------------------------------------

```{r}
rich_meta <- merge(pseq %>% sample_data(), pseq %>% estimate_richness(), by = "row.names")
model_lm <- lm(Shannon ~ Environment, data = rich_meta)
summary(model_lm)
```

Environment after accounting for LP

```{r}
model_lmer <- lmer(Shannon ~ Environment + (1 | LP), data = rich_meta)
summary(model_lmer)
```

------------------------------------------------------------------------

PLAYGROUND

```{r}
df_data <- meta(pseq)
df_data$Shannon <- pseq %>% 
  #rarefy_even_depth() %>% 
  #WrenchWrapper(grp = "Control_type") %>% 
  otu_table() %>% 
  t() %>% 
  diversity(index = "shannon")
  #vegan::specnumber()
fit <- lm(Shannon ~ Environment, data = df_data)
emm <- emmeans(fit, ~ Environment)

# Pairwise comparisons
pairs(emm)

## 
require(pairwiseAdonis)
bray_dist = phyloseq::distance(pseq, method="bray")
p_res <- pairwise.adonis2(x = bray_dist ~ Environment, data = meta(pseq), nperm = 9999)
p_res2 <- pairwise.adonis2(x = bray_dist ~ Environment, data = meta(pseq), strata = "LP", nperm = 9999)
```

### Beta

```{r}
##| fig-width: 16
##| fig-height: 9
##| fig-lightbox: true
pseq_clr <- pseq %>% microbiome::transform(transform = "clr")
#dist_matrix <- phyloseq::distance(pseq_clr, method = "euclidean")
plt1 <- BetaPlot_v02(pseq_clr, strata_f = "Environment", color_f = "LP", shape_f = "Environment", ellipse_f = "Environment", dis_method = "euclidean", title_method = "Hood vs. Bench")
plt2 <- BetaPlot_v02(pseq_clr, strata_f = "LP", color_f = "LP", shape_f = "Environment", ellipse_f = "LP", dis_method = "euclidean", title_method = "Between Laboratory Personnels")
plt <- plt1 + plt2 + plot_layout(guides = "collect")
## save to png
ggsave("./Figures_for_Manuscript_files/saved_png/2.C.svg", plot = plt2, width = 8, height = 6, dpi = 300)
plt2
```

**Figure 2C**: Beta diversity cluster by environmental conditions (left), and technicians (right). Colors depict different technicians, while shapes depict different environmental conditions.

Consider LP with Environment as covariate

Method 1

```{r}
bray_dist <- pseq %>% 
  rarefy_even_depth() %>% 
  phyloseq::distance(method="bray")
clr_euclidean_dist <- pseq %>% 
  microbiome::transform(transform = "clr") %>% 
  phyloseq::distance(method="euclidean")
beta_model_TA <- adonis2(
  bray_dist ~ Environment + LP,
  data = df_data,
  permutations = 9999, 
  by = "terms"
)

beta_model_TA
```

Method 2

```{r}
df_meta <- meta(pseq)
beta_model_block <- adonis2(
  bray_dist ~ LP,
  data = df_meta,
  #permutations = 9999
  permutations = how(
    blocks = df_meta$Environment
  )
)
```

New approach: dbrda with anova

```{r}
Y <- pseq %>% 
  #microbiome::transform(transform = "clr") %>% 
  WrenchWrapper(grp = "LP") %>% 
  abundances() %>% 
  t()
#Y <- decostand(Y, method = "hellinger")
bray_dist = vegan::vegdist(Y, method = "bray") ## euclidean
df_meta <- meta(pseq)
#df_meta$block <- with(df_meta, paste(LP, Environment, sep = "_"))
df_meta$LP <- factor(df_meta$LP)
df_meta$Environment <- factor(df_meta$Environment)
#df_meta$block <- factor(df_meta$block)

#dbrda_mod <- dbrda(bray_dist ~ Control_type + Condition(Person + Year + Quarter), data = df_meta)
dbrda_mod <- dbrda(bray_dist ~ LP, data = df_meta)
results <- anova(dbrda_mod, by = "term", permutations = how(
    blocks = df_meta$Environment)) %>% as.data.frame()
results <- results %>% 
  rownames_to_column(var = "Variable")
```

```{r}
Y <- pseq %>% 
  #WrenchWrapper(grp = "LP") %>% 
  rarefy_even_depth(rngseed = 210488) %>% 
  #microbiome::transform(transform = "clr") %>% 
  abundances() %>% 
  t()
#Y <- decostand(Y, method = "hellinger")
bray_dist = vegan::vegdist(Y, method = "bray")
df_meta <- meta(pseq)
# df_meta$block <- with(df_meta, paste(LP, Environment, sep = "_"))
df_meta$LP <- factor(df_meta$LP)
df_meta$Environment <- factor(df_meta$Environment)
#df_meta$block <- factor(df_meta$block)

#dbrda_mod <- dbrda(bray_dist ~ Control_type + Condition(Person + Year + Quarter), data = df_meta)
dbrda_mod_1 <- dbrda(bray_dist ~ Environment, data = df_meta)
results_1 <- anova(dbrda_mod_1, by = "term", permutations = how(
    blocks = df_meta$LP)) %>% as.data.frame()
results_1 <- results_1 %>% 
  rownames_to_column(var = "Variable") %>% 
  mutate(R2 = SumOfSqs / sum(SumOfSqs)) %>% 
  filter(Variable == "Environment")

dbrda_mod_2 <- dbrda(bray_dist ~ LP, data = df_meta)
results_2 <- anova(dbrda_mod_2, by = "term", permutations = how(
    blocks = df_meta$Environment)) %>% as.data.frame()
results_2 <- results_2 %>% 
  rownames_to_column(var = "Variable") %>% 
  mutate(R2 = SumOfSqs / sum(SumOfSqs)) %>% 
  filter(Variable == "LP")
res <- rbind(results_1, results_2)
```
