---
title: "PDAC technical replica"
author: "AG Neesse - Linh Dang"
editor: visual
execute:
  warning: false
  message: false
format:
  html:
    code-fold: true
    toc: true
    toc-location: left
    toc-depth: 5
    toc-expand: 2
    toc-title: Contents
    fig-responsive: true
    fig-lightbox: true
---

**Library**

```{r message=FALSE}
library(phyloseq)
library(reshape2)
library(microViz)
library(microbiome)
library(vegan)
library(lme4)       # Linear Mixed Models (for repeated measures)
library(lmerTest)   # P-values for Mixed Models
library(emmeans)    # Pairwise comparisons
library(Wrench)
library(rstatix)
library(ggpubr)
library(tidyverse)
library(VennDiagram)
library(ComplexUpset)
rm(list = ls())
```

**Functions**

```{r echo=TRUE}
Kolors <- c("#9d547c","#56ca63","#a357d6","cornflowerblue","#419d2a","sandybrown","red3","peachpuff","cyan","paleturquoise3","mistyrose","mediumpurple","mediumseagreen","mediumorchid","moccasin","orange4","olivedrab","midnightblue","papayawhip","palevioletred4","brown1","greenyellow","orchid","navy","darkred","navajowhite1","mistyrose1","grey85","#525fd6","red2","#8cbe3a","#c944aa","indianred3","#5ba557","#9e66cb","#c1b735","#6d82ec","grey25","#e69728","#6654b0","lightsalmon3","lightcyan1","khaki1","seagreen1","plum1","lightsteelblue1","palevioletred3","mintcream","magenta3","#799330","#da7fdf","#3c782c","#e44586","blue4","#63c996","#dc3f53","#49cbc8","#cf3f29","#4fabda","#da6c2b","#598bd1","#b78c24","#8d4191","#a0b971","slategray1","sienna","plum1","lightyellow1","lightskyblue3","linen","limegreen","cornsilk1","mediumaquamarine","gray14","gold3","darkviolet","#b2386a","#479d71","#ae4341","#2ba198","#e07557","#5361a3","#dda353","#aa98df","#5b6114","#dc89bf","#327243","slateblue1","#e57b94","#277257","#9b62a0","#bbab59","#98495a","#526229","#d8827d","#857624","gray40","#9a4a22","#7c7d46","mediumslateblue","lemonchiffon1","#e3a073","#9e6b33", "gray74","slateblue1","rosybrown3", "lawngreen","gainsboro","dodgerblue3","deeppink3","firebrick3", "orchid2", "olivedrab1", "ivory3", "darkseagreen", "bisque2", "darkgoldenrod2", "blue2", "skyblue", "seashell2", "turquoise", "tan1", "seagreen2", "palevioletred3", "linen", "steelblue4","ghostwhite","dodgerblue1","deeppink1","firebrick1", "limegreen", "purple3", "khaki3", "snow3", "darkslategray","darkorchid","lavender", "magenta2", "palegreen", "salmon", "maroon", "cyan2","#671408","#FAEBD7","#7FFFD4","#F0FFFF","#A52A2A","burlywood","cadetblue","#7FFF00","chocolate","cornsilk","slateblue1","#FF7F50","red1","#008B8B","darkgoldenrod1","darkolivegreen","darkorange4","white","hotpink","honeydew1","goldenrod2","darkgreen","oldlace","darkslategray3","navajowhite3","orchid4","gray25","#F0924D")
WrenchWrapper <- function(PhyloObjct, grp, roundUp = F){
  cnt_table <- PhyloObjct %>% otu_table()
  group <- PhyloObjct %>% sample_data() %>% pull(grp)
  w <- wrench(cnt_table, condition = group)
  
  # deseq.obj <- DESeqDataSetFromMatrix(cnt_table %>% as.data.frame(), DataFrame(group), ~group)
  # DESeq2::sizeFactors(deseq.obj) <- w$nf
  # cnt_table_normalized <- DESeq2::counts(deseq.obj, normalized=TRUE)
  
  norm_factors <- w$nf
  norm_counts <- sweep(cnt_table, 2, norm_factors, FUN = '/')
  if(roundUp){norm_counts <- norm_counts %>% round()}
  if(!is.null(phy_tree(PhyloObjct, errorIfNULL = F))){
    return(phyloseq(otu_table(norm_counts, taxa_are_rows = T), tax_table(PhyloObjct %>% tax_table()), sample_data(PhyloObjct %>% sample_data()),
                    phy_tree(PhyloObjct)))
  } else{
    return(phyloseq(otu_table(norm_counts, taxa_are_rows = T), tax_table(PhyloObjct %>% tax_table()), sample_data(PhyloObjct %>% sample_data())))
  }
}
##
append_AN_NR <- function(pseq, df_additional){
  df_meta <- meta(pseq)
  if(!"AN_NR" %in%  colnames(df_meta)){
    print("There must be AN_NR column!")
    return(NA)
  }
  ## check if info is already in the main data frame, warning if already exits
  col_names_fea <- setdiff(colnames(df_additional), "AN_NR")
  new_col_names_fea <- setdiff(col_names_fea, colnames(df_meta))
  if(length(new_col_names_fea)==0){
    warning("All new feature already exits, nothing to be added")
    return(pseq)
  }
  if(length(new_col_names_fea) < length(col_names_fea)){
    warning("partially info already there, add the rest!")
    df_additional <- df_additional %>% 
      select(c("AN_NR", new_col_names_fea))
  }
  
  df_meta_new <- df_meta %>% 
    rownames_to_column(var = "SampleName") %>% 
    left_join(., df_additional, by = "AN_NR") %>% 
    column_to_rownames(var = "SampleName")
  sample_data(pseq) <- df_meta_new
  return(pseq)
}
## alpha plot function with LMM with fixed and random variables
AlphaPlot_Violin_LMM <- function(df, SampleID, strata, val, metric = "Obsered species"){
  ## df: in pivot longer manner
  ## calc p value in manner of paired
  ## get y.position for ggplot
  y_pos <- t_test(formula = as.formula(sprintf("%s ~ %s", val, strata)), data = df) %>% add_xy_position() %>% pull(y.position)
  model <- lmer(as.formula(sprintf("%s ~ %s + (1|%s)", val, strata, SampleID)), data = df)
  emm <- emmeans(model, specs = as.formula(sprintf("pairwise ~ %s", strata)))
  contr <- contrast(emm, method = "pairwise")
  
  p_add <- contr %>% 
    as_tibble() %>%
    transmute(
      group1 = sub(" - .*", "", contrast),
      group2 = sub(".* - ", "", contrast),
      estimate,
      se = SE,
      statistic = t.ratio,
      df,
      p = p.value
      ##method = "emmeans model-based t-test"
    ) %>% 
    add_significance() %>% 
    mutate(y.position = y_pos)
  plt <- ggplot(df, aes(x = .data[[strata]], y = .data[[val]], fill = .data[[strata]])) +
    geom_violin() +
    #geom_boxplot(alpha = 0.6, outlier.shape = NA) +
    geom_boxplot(width=0.2) +
    geom_point(position = position_jitter(width = 0.1), alpha = 0.5) +
    # Connect lines for the same sample to show paired changes
    geom_line(aes(group = AN_NR), color = "gray", alpha = 0.5) +
    theme_minimal() +
    labs(
      title = "Comparison of Decontamination Methods",
      subtitle = sprintf("Alpha Diversity: %s", metric),
      y = metric
    ) +
    theme(legend.position = "none")
  plt <- plt + stat_pvalue_manual(p_add, label = "p.signif", inherit.aes = FALSE, tip.length = 0.01)
  return(list(plt = plt, const = contr))
}

## alpha plot with Wilcon_test
AlphaPlot_Violin_Wilcox <- function(df, SampleID, strata, val, y_label = "Obsered species", subtitle = "The higher the better", ...){
  ## df: in pivot longer manner
  ## calc p value in manner of paired
  ## get y.position for ggplot
  p_add <- wilcox_test(formula = as.formula(sprintf("%s ~ %s", val, strata)), data = df, paired = TRUE, ...) %>% add_significance() %>%  add_xy_position() ##%>% pull(y.position)
  
  plt <- ggplot(df, aes(x = .data[[strata]], y = .data[[val]], fill = .data[[strata]])) +
    geom_violin() +
    #geom_boxplot(alpha = 0.6, outlier.shape = NA) +
    geom_boxplot(width=0.2) +
    geom_point(position = position_jitter(width = 0.1), alpha = 0.5) +
    # Connect lines for the same sample to show paired changes
    geom_line(aes(group = AN_NR), color = "gray", alpha = 0.5) +
    theme_minimal() +
    labs(
      title = "Paired Wilcox test",
      subtitle = subtitle,
      y = y_label
    ) +
    theme(legend.position = "none")
  plt <- plt + stat_pvalue_manual(p_add, label = "p.adj.signif", inherit.aes = FALSE, tip.length = 0.01)
  return(plt)
}
##------------
set_size = function(w, h, factor=1.5) {
    s = 1 * factor
    options(
        repr.plot.width=w * s,
        repr.plot.height=h * s,
        repr.plot.res=100 / factor,
        jupyter.plot_mimetypes='image/png',
        jupyter.plot_scale=1
    )
}
```

**Load original data**

```{r}
#| include: false
## Load data
#source("../../../add_code/Functions.R")
df_additionalInfo <- readxl::read_xlsx("../../../meta/Mice_meta.xlsx")
pseq_raw <- readRDS("../../../data/Chap3/pseq_Proj5_postFilter_v04.rds") %>% 
  ps_filter(ffpe.bulk == "bulk") %>% 
  ps_filter(true.control == "true") %>% 
  append_AN_NR(df_additional = df_additionalInfo) #%>% 
  #WrenchWrapper(grp = "Sex") 
pseq_restrictive <- readRDS("../../../data/Chap3/pseq_bulk_restrictive.rds") %>% 
  append_AN_NR(df_additionalInfo) #%>% 
  #WrenchWrapper(grp = "Sex")
pseq_decontam <- readRDS("../../../data/Chap3/pseq_bulk_decontam_p0.5.rds") %>% 
  append_AN_NR(df_additionalInfo) #%>% 
  #WrenchWrapper(grp = "Sex")
pseq_SCRuB <- readRDS("../../../data/Chap3/pseq_bulk_SCRuB.rds") %>% 
  append_AN_NR(df_additionalInfo) #%>% 
  #WrenchWrapper(grp = "Sex")
pseq_Nj <- readRDS("../../../data/Chap3/pseq_bulk_Fisher_v02.rds") %>% 
  append_AN_NR(df_additionalInfo) #%>% 
  #WrenchWrapper(grp = "Sex")
```

**Load replica**

```{r}
rep_raw <- readRDS("../../../data/Chap3_Addition/pseq_origin_v1.0.rds") %>% 
  ps_filter(ffpe.bulk == "bulk") %>% 
  append_AN_NR(df_additional = df_additionalInfo)
rep_restrictive <- readRDS("../../../data/Chap3_Addition/pseq_restrictive.rds") %>% 
  append_AN_NR(df_additional = df_additionalInfo)
rep_decontam <- readRDS("../../../data/Chap3_Addition/pseq_decontam.rds") %>% 
  append_AN_NR(df_additional = df_additionalInfo)
rep_SCRuB <- readRDS("../../../data/Chap3_Addition/pseq_SCRuB.rds") %>% 
  append_AN_NR(df_additional = df_additionalInfo)
rep_Nj <- readRDS("../../../data/Chap3_Addition/pseq_Nejman.rds") %>% 
  append_AN_NR(df_additional = df_additionalInfo)
## rename taxa
taxa_names(rep_raw) <- gsub(pattern = " ", replacement = "_", x = taxa_names(rep_raw))
taxa_names(rep_restrictive) <- gsub(pattern = " ", replacement = "_", x = taxa_names(rep_restrictive))
taxa_names(rep_decontam) <- gsub(pattern = " ", replacement = "_", x = taxa_names(rep_decontam))
taxa_names(rep_SCRuB) <- gsub(pattern = " ", replacement = "_", x = taxa_names(rep_SCRuB))
taxa_names(rep_Nj) <- gsub(pattern = " ", replacement = "_", x = taxa_names(rep_Nj))
```

Retain only sample with pair

```{r}
IDs <- meta(rep_raw)[["AN_NR"]]
pseq_raw <- pseq_raw %>% ps_filter(AN_NR %in% IDs)
pseq_restrictive <- pseq_restrictive %>% ps_filter(AN_NR %in% IDs)
pseq_decontam <- pseq_decontam %>% ps_filter(AN_NR %in% IDs)
pseq_SCRuB <- pseq_SCRuB %>% ps_filter(AN_NR %in% IDs)
pseq_Nj <- pseq_Nj %>% ps_filter(AN_NR %in% IDs)
```

# Project Structure through Mind-Map

```{mermaid}
flowchart TD
  %%A[Hard edge] --> B(Round edge)
  %%A --> F(Others)
  %%B --> C{Decision}
  %%C --> D[Result one]
  %%C --> E[Result two]
  
  S(Replica Analysis) --> B(Properties of Replica)
  S --> C(Overlap)
  C --> C1(UpsetPlot-Phylum)
  C --> C2(UpsetPlot-Class)
  S --> D(Assessment)
  D --> D1(1.Intra-Sample Stability)
  D --> D2(2.PERANOVA)
  
```

# Replica Properties

## Alpha

Diversity between

## Beta

# Assessment

Using replica as additional dataset for assessment of decontamination approaches.

## 1. Intra-samples Stability

-   **Assumption**: after removing contaminants, bacterial profile of a sample should be stable, regardless of sequencing year or technician.
-   **Method**: using appropriate distance function to measure the (dis)similarity of the sample sample w.r.t 2 technical replicas. Afterwards, applying paired t-test or Wilcoxon Signed-Rank Test, LMM, or constrained permutation test.

### 1.1 Bray-curtis distance with compositional normalization

```{r}
## Write a wrapper

process_replica <- function(
    ps1, ## original sample
    ps2, ## replica 
    normalize = "compositional", ## compositional - bray | clr - euclidean
    distance = "bray",
    col_name = "raw"){
  ## add replica info and change names
  ps1 <- ps1 %>% 
    ps_mutate(Replica = "R1") %>% 
    ps_mutate(NewName = sprintf("%s_%s", AN_NR, Replica))
  sample_names(ps1) <- meta(ps1)[["NewName"]]
  
  ps2 <- ps2 %>% 
    ps_mutate(Replica = "R2") %>% 
    ps_mutate(NewName = sprintf("%s_%s", AN_NR, Replica))
  sample_names(ps2) <- meta(ps2)[["NewName"]]
  
  ps_merged <- merge_phyloseq(ps1, ps2)
  if(normalize == "rarefaction"){
    ps_merged <- ps_merged %>% rarefy_even_depth()
  }else if(normalize == "Wrench"){
    ps_merged <- ps_merged %>% WrenchWrapper(grp = "Sex", roundUp = FALSE)
  } else if(normalize %in%  c("clr", "compositional")){
    ps_merged <- ps_merged %>% microbiome::transform(transform = normalize)
  }
  
  # merge_phyloseq(ps1, ps2) %>% 
  #   microbiome::transform(transform = normalize) %>% 
  dis_mx <- NULL
  if(distance == "jaccard"){
    dis_mx <- ps_merged %>% 
      phyloseq::distance(method = distance, binary = TRUE)
  } else{
    dis_mx <- ps_merged %>% 
      phyloseq::distance(method = distance)
  }
  dis_mx <- dis_mx %>% 
    # ps_merged %>% 
    # phyloseq::distance(method = distance) %>% 
    as.matrix() %>% 
    as.data.frame() %>% 
    dplyr::select(contains("R2")) %>% 
    rownames_to_column(var = "rowname") %>% 
    filter(grepl("R1", rowname)) %>% 
    arrange(rowname) %>% 
    column_to_rownames("rowname") %>% 
    select(sort(names(.)))
  col <- tibble(AN_NR = sapply(rownames(dis_mx), FUN = function(x){strsplit(x, split = "_")[[1]][1]}), !!sym(col_name) := diag(as.matrix(dis_mx)))
  return(col)
}
```

Create wide-format table

```{r}
df <- process_replica(pseq_raw, rep_raw, col_name = "Raw") %>% 
  merge(., process_replica(pseq_restrictive, rep_restrictive, col_name = "Restrictive"), by = "AN_NR") %>%
  merge(., process_replica(pseq_decontam, rep_decontam, col_name = "Decontam"), by="AN_NR") %>% 
  merge(., process_replica(pseq_SCRuB, rep_SCRuB, col_name = "SCRuB"), by="AN_NR") %>% 
  merge(., process_replica(pseq_Nj, rep_Nj, col_name = "Nj"), by="AN_NR")
df_longer <- df %>% 
  pivot_longer(cols = -AN_NR, names_to = "Methods", values_to = "Val")
#df_longer_1 <- df_longer %>% filter(Methods %in% c("Raw", "Nj"))
print(df)
```

Wilcox-test on Distance

```{r}
df_longer$Methods <- factor(df_longer$Methods, levels = c("Raw", "Restrictive", "Decontam", "SCRuB", "Nj"))
AlphaPlot_Violin_Wilcox(df_longer, "AN_NR", "Methods", "Val", y_label = "bray distance", subtitle = "The lower the better", comparisons = list(c("Raw", "Restrictive"), c("Raw", "Decontam"), c("Raw", "SCRuB"), c("Raw", "Nj")), alternative = "greater")
```

Negative Log Distance (only for a bit nicer illustration)

```{r}
df_longer %>% 
  mutate(Val_NegLog = -log(Val)) %>% 
AlphaPlot_Violin_Wilcox("AN_NR", "Methods", "Val_NegLog", y_label = "Neg Log b-ray distance", subtitle = "The higher the better", comparisons = list(c("Raw", "Restrictive"), c("Raw", "Decontam"), c("Raw", "SCRuB"), c("Raw", "Nj")), alternative = "less")
```

**Remarks**: Decontam and Nj seems to have smaller distance among replicates.

#### For each sample, plot beta plot

```{r}
# edit metadata
# pseq_raw <- pseq_raw %>% 
#   ps_mutate(Method = "Raw", Replica = "R1") %>% 
#   ps_mutate(NewName = sprintf("%s_%s_%s", AN_NR, Method, Replica))
# sample_names(pseq_raw) <- meta(pseq_raw)[["NewName"]]
# pseq_raw <- pseq_raw %>% ps_select(-NewName)
# 
# lapply()

lst_ps <- list(Raw = pseq_raw, Restrictive = pseq_restrictive, Decontam = pseq_decontam, SCRuB = pseq_SCRuB, Nj = pseq_Nj)
lst_rep <- list(Raw = rep_raw, Restrictive = rep_restrictive, Decontam = rep_decontam, SCRuB = rep_SCRuB, Nj = rep_Nj)

lst_ps <- lapply(seq_along(lst_ps), function(i){
  nm <- names(lst_ps[i])
  res <- lst_ps[[i]] %>% 
    ps_mutate(Method = nm, Replica = "R1") %>% 
    ps_mutate(NewName = sprintf("%s_%s_%s", AN_NR, Method, Replica))
  sample_names(res) <- meta(res)[["NewName"]]
  res <- res %>% ps_select(-NewName)
  return(res)
}) %>% unlist()
names(lst_ps) <- c("Raw", "Restrictive", "Decontam", "SCRuB", "Nj")
lst_rep <- lapply(seq_along(lst_rep), function(i){
  nm <- names(lst_rep[i])
  res <- lst_rep[[i]] %>% 
    ps_mutate(Method = nm, Replica = "R2") %>% 
    ps_mutate(NewName = sprintf("%s_%s_%s", AN_NR, Method, Replica))
  sample_names(res) <- meta(res)[["NewName"]]
  res <- res %>% ps_select(-NewName)
  return(res)
}) %>% unlist()
names(lst_rep) <- c("Raw", "Restrictive", "Decontam", "SCRuB", "Nj")
#id <- "AN1023"

# ps_1023 <- merge_phyloseq(lst_ps$Raw, lst_ps$Restrictive, lst_ps$Decontam, lst_ps$SCRuB, lst_ps$Nj,
#                           lst_rep$Raw, lst_rep$Restrictive, lst_rep$Decontam, lst_rep$SCRuB, lst_rep$Nj)
# ps_1023 <- ps_1023 %>% prune_samples(meta(.)[["AN_NR"]] == "AN1023", .)
# ord <- ps_1023 %>% microbiome::transform(transform = "identity") %>% 
#   ordinate(method = "PCoA", distance = "jaccard")
# 
# plot_ordination(ps_1023, ord, color = "Method", shape = "Replica") +
#   geom_point(size = 3) +
#   theme_bw() +
#   labs(title = "Beta diversity (Bray–Curtis, PCoA)")


samplewise_plot <- function(pseq_merge, Sample_ID, normalized = "identity", dist_metric = "bray"){
  ## normalized: compositional, rarefaction, Wrench
  ## dist_metric: bray, unifrac, 
  pseq_one <- pseq_merge %>% 
    prune_samples(meta(.)[["AN_NR"]] == Sample_ID, .) %>% 
    prune_taxa(taxa_sums(.) >0, .)
  if(normalized == "rarefaction"){
    pseq_one <- pseq_one %>% rarefy_even_depth()
  }else if(normalized == "Wrench"){
    pseq_one <- pseq_one %>% WrenchWrapper(grp = "Sex", roundUp = FALSE)
  } else if(normalized %in%  c("clr", "compositional", "identity")){
    pseq_one <- pseq_one %>% microbiome::transform(transform = normalized)
  }
  beta_dis = NULL
  if(dist_metric %in%  c("bray", "jaccard", "unifrac", "jsd", "euclidean")){
    beta_dis <- pseq_one %>% phyloseq::distance(method = dist_metric)
  }
  # print(pseq_one)
  # print(pseq_one %>% meta())
  # print(beta_dis)
  ord <- phyloseq::ordinate(pseq_one, method = "PCoA", distance = beta_dis)
  # print(ord)
  plt <- plot_ordination(pseq_one, ord, color = "Method", shape = "Replica") +
  geom_point(size = 4, alpha = 0.8) +
  theme_bw() +
  ##stat_ellipse(aes(group = Condition), linetype = 2) + # Add confidence ellipses
  labs(title = sprintf("%s: %s Distance - %s transform", Sample_ID, dist_metric, normalized),
       subtitle = sprintf("Colored by Methods, Shaped by Replica"))
  return(plt)
} 
```

Plot

```{r}
pseq_merge <- merge_phyloseq(lst_ps$Raw, lst_ps$Restrictive, lst_ps$Decontam, lst_ps$SCRuB, lst_ps$Nj,
                             lst_rep$Raw, lst_rep$Restrictive, lst_rep$Decontam, lst_rep$SCRuB, lst_rep$Nj)
IDs <- meta(pseq_raw)[["AN_NR"]]
plt <- sapply(IDs, FUN = function(x){samplewise_plot(pseq_merge, x, normalized = "compositional", dist_metric = "bray")}) 
require(patchwork)
plt[1]
```

```{r}
plt[2]
```

```{r}
plt[3]
```

```{r}
plt[4]
```

```{r}
plt[5]
```

```{r}
plt[6]
```

```{r}
plt[7]
```

```{r}
plt[8]
```

```{r}
plt[9]
```

```{r}
plt[10]
```

### 1.2 Bray-curtis distance with rarefaction

Remove AN986 due to low read count.

```{r}
df <- process_replica(pseq_raw %>% ps_filter(AN_NR != "AN986"), rep_raw %>%  ps_filter(AN_NR != "AN986"), normalize = "rarefaction", distance = "bray", col_name = "Raw") %>% 
  left_join(., process_replica(pseq_restrictive %>% ps_filter(AN_NR != "AN986"), rep_restrictive %>% ps_filter(AN_NR != "AN986"), normalize = "rarefaction", distance = "bray", col_name = "Restrictive"), by = "AN_NR") %>%
  left_join(., process_replica(pseq_decontam %>% ps_filter(AN_NR != "AN986"), rep_decontam %>% ps_filter(AN_NR != "AN986"), normalize = "rarefaction", distance = "bray", col_name = "Decontam"), by="AN_NR") %>% 
  left_join(., process_replica(pseq_SCRuB %>% ps_filter(AN_NR != "AN986"), rep_SCRuB %>% ps_filter(AN_NR != "AN986"), normalize = "rarefaction", distance = "bray", col_name = "SCRuB"), by="AN_NR") %>% 
  left_join(., process_replica(pseq_Nj %>% ps_filter(AN_NR != "AN986"), rep_Nj %>% ps_filter(AN_NR != "AN986"), normalize = "rarefaction", distance = "bray", col_name = "Nj"), by="AN_NR")
df_longer <- df %>% 
  pivot_longer(cols = -AN_NR, names_to = "Methods", values_to = "Val")
#df_longer_1 <- df_longer %>% filter(Methods %in% c("Raw", "Nj"))
print(df)
```

Wilcox-test

```{r}
AlphaPlot_Violin_Wilcox(df_longer, "AN_NR", "Methods", "Val", y_label = "b-ray distance", subtitle = "The lower the better", comparisons = list(c("Raw", "Restrictive"), c("Raw", "Decontam"), c("Raw", "SCRuB"), c("Raw", "Nj")), alternative = "greater")
```

```{r}
plt <- sapply(setdiff(IDs, "AN986"), FUN = function(x){samplewise_plot(pseq_merge, x, normalized = "rarefaction", dist_metric = "bray")}) 
```

### 1.2+ Bray-curtis distance with Wrench normalization

```{r}
dist_metric <- "bray"
trans <-"Wrench"  #"rarefaction"
df <- process_replica(pseq_raw, rep_raw, normalize = trans, distance = dist_metric, col_name = "Raw") %>% 
  left_join(., process_replica(pseq_restrictive, rep_restrictive, normalize = trans, distance = dist_metric, col_name = "Restrictive"), by = "AN_NR") %>%
  left_join(., process_replica(pseq_decontam, rep_decontam, normalize = trans, distance = dist_metric, col_name = "Decontam"), by="AN_NR") %>% 
  left_join(., process_replica(pseq_SCRuB, rep_SCRuB, normalize = trans, distance = dist_metric, col_name = "SCRuB"), by="AN_NR") %>% 
  left_join(., process_replica(pseq_Nj, rep_Nj, normalize = trans, distance = dist_metric, col_name = "Nj"), by="AN_NR")
df_longer <- df %>% 
  pivot_longer(cols = -AN_NR, names_to = "Methods", values_to = "Val")
#df_longer_1 <- df_longer %>% filter(Methods %in% c("Raw", "Nj"))
print(df)
```

Wilcox test

```{r}
df_longer$Methods <- factor(df_longer$Methods, levels = c("Raw", "Restrictive", "Decontam", "SCRuB", "Nj"))
res <- wilcox_test(data = df_longer, formula = Val ~ Methods, paired = TRUE, comparisons = list(c("Raw", "Restrictive"), c("Raw", "Decontam"), c("Raw", "SCRuB"), c("Raw", "Nj")), alternative = "greater")
print(res)
```

### 1.3 Jarcard distance with rarefaction

Remove AN986 due to low read count.

```{r}
dist_metric <- "jaccard"
trans <-"identity"  #"rarefaction"
df <- process_replica(pseq_raw %>% ps_filter(AN_NR != "AN986"), rep_raw %>%  ps_filter(AN_NR != "AN986"), normalize = trans, distance = dist_metric, col_name = "Raw") %>% 
  left_join(., process_replica(pseq_restrictive %>% ps_filter(AN_NR != "AN986"), rep_restrictive %>% ps_filter(AN_NR != "AN986"), normalize = trans, distance = dist_metric, col_name = "Restrictive"), by = "AN_NR") %>%
  left_join(., process_replica(pseq_decontam %>% ps_filter(AN_NR != "AN986"), rep_decontam %>% ps_filter(AN_NR != "AN986"), normalize = trans, distance = dist_metric, col_name = "Decontam"), by="AN_NR") %>% 
  left_join(., process_replica(pseq_SCRuB %>% ps_filter(AN_NR != "AN986"), rep_SCRuB %>% ps_filter(AN_NR != "AN986"), normalize = trans, distance = dist_metric, col_name = "SCRuB"), by="AN_NR") %>% 
  left_join(., process_replica(pseq_Nj %>% ps_filter(AN_NR != "AN986"), rep_Nj %>% ps_filter(AN_NR != "AN986"), normalize = trans, distance = dist_metric, col_name = "Nj"), by="AN_NR")
df_longer <- df %>% 
  pivot_longer(cols = -AN_NR, names_to = "Methods", values_to = "Val")
#df_longer_1 <- df_longer %>% filter(Methods %in% c("Raw", "Nj"))
print(df)
```

Wilcox

```{r}
df_longer$Methods <- factor(df_longer$Methods, levels = c("Raw", "Restrictive", "Decontam", "SCRuB", "Nj"))
AlphaPlot_Violin_Wilcox(df_longer, "AN_NR", "Methods", "Val", y_label = "Jarcard distance", subtitle = "The lower the better", comparisons = list(c("Raw", "Restrictive"), c("Raw", "Decontam"), c("Raw", "SCRuB"), c("Raw", "Nj")), alternative = "greater")
```

For a nicer view,

```{r}
df_longer %>% 
  mutate(Val_NegLog = -1*log(Val)) %>% 
  AlphaPlot_Violin_Wilcox("AN_NR", "Methods", "Val_NegLog", y_label = "Jarcard distance", subtitle = "The higher the better", comparisons = list(c("Raw", "Restrictive"), c("Raw", "Decontam"), c("Raw", "SCRuB"), c("Raw", "Nj")), alternative = "less")
```

### 1.3+ Jarcard distance with Wrench normalization

```{r}
dist_metric <- "jaccard"
trans <-"Wrench"  #"rarefaction"
df <- process_replica(pseq_raw, rep_raw, normalize = trans, distance = dist_metric, col_name = "Raw") %>% 
  left_join(., process_replica(pseq_restrictive, rep_restrictive, normalize = trans, distance = dist_metric, col_name = "Restrictive"), by = "AN_NR") %>%
  left_join(., process_replica(pseq_decontam, rep_decontam, normalize = trans, distance = dist_metric, col_name = "Decontam"), by="AN_NR") %>% 
  left_join(., process_replica(pseq_SCRuB, rep_SCRuB, normalize = trans, distance = dist_metric, col_name = "SCRuB"), by="AN_NR") %>% 
  left_join(., process_replica(pseq_Nj, rep_Nj, normalize = trans, distance = dist_metric, col_name = "Nj"), by="AN_NR")
df_longer <- df %>% 
  pivot_longer(cols = -AN_NR, names_to = "Methods", values_to = "Val")
#df_longer_1 <- df_longer %>% filter(Methods %in% c("Raw", "Nj"))
print(df)
```

Wilcox

```{r}
df_longer$Methods <- factor(df_longer$Methods, levels = c("Raw", "Restrictive", "Decontam", "SCRuB", "Nj"))
AlphaPlot_Violin_Wilcox(df_longer, "AN_NR", "Methods", "Val", y_label = "Jarcard distance", subtitle = "The lower the better", comparisons = list(c("Raw", "Restrictive"), c("Raw", "Decontam"), c("Raw", "SCRuB"), c("Raw", "Nj")), alternative = "greater")
```

### 1.4 **Aitchison - CLR transform with eucledian distance**

```{r}
dist_metric <- "euclidean"
trans <-"clr"  #"rarefaction"
df <- process_replica(pseq_raw, rep_raw, normalize = trans, distance = dist_metric, col_name = "Raw") %>% 
  left_join(., process_replica(pseq_restrictive, rep_restrictive, normalize = trans, distance = dist_metric, col_name = "Restrictive"), by = "AN_NR") %>%
  left_join(., process_replica(pseq_decontam, rep_decontam, normalize = trans, distance = dist_metric, col_name = "Decontam"), by="AN_NR") %>% 
  left_join(., process_replica(pseq_SCRuB, rep_SCRuB, normalize = trans, distance = dist_metric, col_name = "SCRuB"), by="AN_NR") %>% 
  left_join(., process_replica(pseq_Nj, rep_Nj, normalize = trans, distance = dist_metric, col_name = "Nj"), by="AN_NR")
df_longer <- df %>% 
  pivot_longer(cols = -AN_NR, names_to = "Methods", values_to = "Val")
#df_longer_1 <- df_longer %>% filter(Methods %in% c("Raw", "Nj"))
print(df)
```

Wilcox

```{r}
df_longer$Methods <- factor(df_longer$Methods, levels = c("Raw", "Restrictive", "Decontam", "SCRuB", "Nj"))
AlphaPlot_Violin_Wilcox(df_longer, "AN_NR", "Methods", "Val", y_label = "Aitchison distance", subtitle = "The lower the better", alternative = "greater")
```

Plot for each sample

```{r}
plt <- sapply(IDs, FUN = function(x){samplewise_plot(pseq_merge, x, normalized = "clr", dist_metric = "euclidean")}) 
```

```{r}
plt[1]
```

```{r}
plt[2]
```

```{r}
plt[3]
```

```{r}
plt[4]
```

```{r}
plt[5]
```

```{r}
plt[6]
```

```{r}
plt[7]
```

```{r}
plt[8]
```

```{r}
plt[9]
```

```{r}
plt[10]
```

**Remarks**: all decontamination methods seem to improve the data. Among them, Nj outperformed others.

## 2. PERANOVA based Approach

Linear Mixed Model: distance \~ AN_NR + Reaplica

Assumption: With high quality data, the different among samples should be explain by bacterial profile of samples, instead of replica.

<!-- Each Method, calculate distance and adonis2 using dis ~ AN_NR + batch and compare R2. -->

```{r}
process_pernova <- function(ps1, ps2, transform = "compositional", dist_metric = "bray"){
  ps1 <- ps1 %>% 
    ps_mutate(Replica = "R1") %>% 
    ps_mutate(NewName = sprintf("%s_%s", AN_NR, Replica))
  sample_names(ps1) <- meta(ps1)[["NewName"]]
    
  ps2 <- ps2 %>% 
    ps_mutate(Replica = "R2") %>% 
    ps_mutate(NewName = sprintf("%s_%s", AN_NR, Replica))
  sample_names(ps2) <- meta(ps2)[["NewName"]]
  
  ps <- merge_phyloseq(ps1, ps2)
  #print(meta(ps))
  dis_mx <- ps %>% 
    microbiome::transform(transform = transform) %>% 
    phyloseq::distance(method = dist_metric)
  
  res <- vegan::adonis2(formula = dis_mx ~ AN_NR + Replica, data = meta(ps), by = "margin", permutations = 99999)
  return(res)
}
```

Compare among different approaches

```{r}
df_res <- data.frame(Method = character(0), AN_NR = double(0), Rep = double(0))
df_res[nrow(df_res) +1, ] <- c("Raw",process_pernova(pseq_raw, rep_raw, transform = "clr", dist_metric = "euclidean")$R2[1:2] %>% unlist())
df_res[nrow(df_res) +1, ] <- c("Restrictive",process_pernova(pseq_restrictive, rep_restrictive, transform = "clr", dist_metric = "euclidean")$R2[1:2] %>% unlist())
df_res[nrow(df_res) +1, ] <- c("Decontam",process_pernova(pseq_decontam, rep_decontam, transform = "clr", dist_metric = "euclidean")$R2[1:2] %>% unlist())
df_res[nrow(df_res) +1, ] <- c("SCRuB",process_pernova(pseq_SCRuB, rep_SCRuB, transform = "clr", dist_metric = "euclidean")$R2[1:2] %>% unlist())
df_res[nrow(df_res) +1, ] <- c("Nj",process_pernova(pseq_Nj, rep_Nj, transform = "clr", dist_metric = "euclidean")$R2[1:2] %>% unlist())
print(df_res)
```

Remarks: SCRuB outperforms the other

## 3. Overlap

Investigate the microbial profile of 2 technical replica

Facet by Replica

```{r}
## create df_R1
df_R1 <- data.frame(OTU = taxa_names(pseq_raw), Raw = 1) %>% 
  left_join(., data.frame(OTU = taxa_names(pseq_restrictive), Restrictive = 1), by = "OTU") %>% 
  left_join(., data.frame(OTU = taxa_names(pseq_decontam), Decontam = 1), by = "OTU") %>% 
  left_join(., data.frame(OTU = taxa_names(pseq_SCRuB), SCRuB = 1), by = "OTU") %>% 
  left_join(., data.frame(OTU = taxa_names(pseq_Nj), Nj = 1), by = "OTU") %>% 
  select(-Raw) %>% 
  dplyr::mutate_if(is.numeric, coalesce, 0)
df_R1_tax <- tax_tibble(pseq_raw)
df_R1 <- df_R1 %>% 
  left_join(., df_R1_tax, by = c("OTU" = "FeatureID"))

## create df_R2
df_R2 <- data.frame(OTU = taxa_names(rep_raw), Raw = 1) %>% 
  left_join(., data.frame(OTU = taxa_names(rep_restrictive), Restrictive = 1), by = "OTU") %>% 
  left_join(., data.frame(OTU = taxa_names(rep_decontam), Decontam = 1), by = "OTU") %>% 
  left_join(., data.frame(OTU = taxa_names(rep_SCRuB), SCRuB = 1), by = "OTU") %>% 
  left_join(., data.frame(OTU = taxa_names(rep_Nj), Nj = 1), by = "OTU") %>% 
  select(-Raw) %>% 
  dplyr::mutate_if(is.numeric, coalesce, 0)
df_R2_tax <- tax_tibble(rep_raw)
df_R2 <- df_R2 %>% 
  left_join(., df_R2_tax, by = c("OTU" = "FeatureID"))

  #dplyr::mutate_if(is.numeric, coalesce, 0)
```

Phylum

```{r}
#| fig-width: 14
#| fig-height: 10
#| fig-lightbox: true
require(patchwork)
manual_fill = c('Actinomycetota'=Kolors[1], 'Bacillota'=Kolors[2], 'Bacteroidota'= Kolors[3], 'Campylobacterota'=Kolors[4], 'Cyanobacteriota'=Kolors[5], 'Pseudomonadota'=Kolors[6], 'Thermodesulfobacteriota'=Kolors[7], 'Other'=Kolors[8])
df_R1 <- df_R1 %>% mutate(Phylum = if_else(phylum %in% names(manual_fill), phylum, "Other"))
df_R2 <- df_R2 %>% mutate(Phylum = if_else(phylum %in% names(manual_fill), phylum, "Other"))

Medthods <- c("Restrictive", "Decontam", "SCRuB", "Nj")
annotations = list(
    'Phylum' = list(
      aes = aes(x=intersection, fill=Phylum),
      geom = list(
        geom_bar(stat='count', position='fill', na.rm=TRUE),
        # geom_text(
        #   aes(label=ifelse(phylum == 'Pseudomonadota', 'Pseu.', NA)),
        #   stat='count',
        #   position=position_fill(vjust = .5),
        #   na.rm=TRUE
        # ),
        # scale_color_manual(values=c('show'='black', 'hide'='transparent'), guide='none'),
        scale_fill_manual(values = manual_fill)
        
      )
    )
)
set.seed(0)    # for replicable example only

u1 <- upset(df_R1, Medthods, annotations =  annotations, name='R1', width_ratio=0.1)
u2 <- upset(df_R2, Medthods, annotations =  annotations, name='R2', width_ratio=0.1)
(u1 | u2) + plot_layout(guides='collect')
```

Class

```{r}
#| fig-width: 14
#| fig-height: 8
#| fig-lightbox: true
#set_size(16, 5)
manual_fill_class <- c("Alphaproteobacteria"=Kolors[1], "Bacilli"=Kolors[2], "Clostridia"=Kolors[3], "Gammaproteobacteria"=Kolors[4], "Bacteroidia"=Kolors[5], "Betaproteobacteria"=Kolors[6], "Actinomycetes"=Kolors[7], "Flavobacteriia"=Kolors[8], "Negativicutes"=Kolors[9], "Cyanophyceae"=Kolors[10], "Other"=Kolors[11])
df_R1 <- df_R1 %>% mutate(Class = if_else(class %in% names(manual_fill_class), class, "Other"))
df_R2 <- df_R2 %>% mutate(Class = if_else(class %in% names(manual_fill_class), class, "Other"))

annotations_class = list(
    'Class' = list(
      aes = aes(x=intersection, fill=Class),
      geom = list(
        geom_bar(stat='count', position='fill', na.rm=TRUE),
        # geom_text(
        #   aes(label=ifelse(phylum == 'Pseudomonadota', 'Pseu.', NA)),
        #   stat='count',
        #   position=position_fill(vjust = .5),
        #   na.rm=TRUE
        # ),
        # scale_color_manual(values=c('show'='black', 'hide'='transparent'), guide='none'),
        scale_fill_manual(values = manual_fill_class)
        
      )
    )
)
set.seed(0)    # for replicable example only

u1 <- upset(df_R1, Medthods, annotations =  annotations_class, name='R1', width_ratio=0.1)
u2 <- upset(df_R2, Medthods, annotations =  annotations_class, name='R2', width_ratio=0.1)

(u1 | u2) + plot_layout(guides='collect')
```
