---
title: "PDAC technical replica"
author: "AG Neesse - Linh Dang"
editor: visual
execute:
  warning: false
  message: false
format:
  html:
    code-fold: true
    toc: true
    toc-location: left
    toc-depth: 3
    fig-responsive: true
    fig-lightbox: true
---

```{r message=FALSE}
library(phyloseq)
library(reshape2)
library(microViz)
library(microbiome)
library(vegan)
library(lme4)       # Linear Mixed Models (for repeated measures)
library(lmerTest)   # P-values for Mixed Models
library(emmeans)    # Pairwise comparisons
library(Wrench)
library(rstatix)
library(ggpubr)
library(tidyverse)
library(VennDiagram)
library(ComplexUpset)
rm(list = ls())
```

**Functions**

```{r echo=TRUE}
Kolors <- c("#9d547c","#56ca63","#a357d6","cornflowerblue","#419d2a","sandybrown","red3","peachpuff","cyan","paleturquoise3","mistyrose","mediumpurple","mediumseagreen","mediumorchid","moccasin","orange4","olivedrab","midnightblue","papayawhip","palevioletred4","brown1","greenyellow","orchid","navy","darkred","navajowhite1","mistyrose1","grey85","#525fd6","red2","#8cbe3a","#c944aa","indianred3","#5ba557","#9e66cb","#c1b735","#6d82ec","grey25","#e69728","#6654b0","lightsalmon3","lightcyan1","khaki1","seagreen1","plum1","lightsteelblue1","palevioletred3","mintcream","magenta3","#799330","#da7fdf","#3c782c","#e44586","blue4","#63c996","#dc3f53","#49cbc8","#cf3f29","#4fabda","#da6c2b","#598bd1","#b78c24","#8d4191","#a0b971","slategray1","sienna","plum1","lightyellow1","lightskyblue3","linen","limegreen","cornsilk1","mediumaquamarine","gray14","gold3","darkviolet","#b2386a","#479d71","#ae4341","#2ba198","#e07557","#5361a3","#dda353","#aa98df","#5b6114","#dc89bf","#327243","slateblue1","#e57b94","#277257","#9b62a0","#bbab59","#98495a","#526229","#d8827d","#857624","gray40","#9a4a22","#7c7d46","mediumslateblue","lemonchiffon1","#e3a073","#9e6b33", "gray74","slateblue1","rosybrown3", "lawngreen","gainsboro","dodgerblue3","deeppink3","firebrick3", "orchid2", "olivedrab1", "ivory3", "darkseagreen", "bisque2", "darkgoldenrod2", "blue2", "skyblue", "seashell2", "turquoise", "tan1", "seagreen2", "palevioletred3", "linen", "steelblue4","ghostwhite","dodgerblue1","deeppink1","firebrick1", "limegreen", "purple3", "khaki3", "snow3", "darkslategray","darkorchid","lavender", "magenta2", "palegreen", "salmon", "maroon", "cyan2","#671408","#FAEBD7","#7FFFD4","#F0FFFF","#A52A2A","burlywood","cadetblue","#7FFF00","chocolate","cornsilk","slateblue1","#FF7F50","red1","#008B8B","darkgoldenrod1","darkolivegreen","darkorange4","white","hotpink","honeydew1","goldenrod2","darkgreen","oldlace","darkslategray3","navajowhite3","orchid4","gray25","#F0924D")
WrenchWrapper <- function(PhyloObjct, grp, roundUp = F){
  cnt_table <- PhyloObjct %>% otu_table()
  group <- PhyloObjct %>% sample_data() %>% pull(grp)
  w <- wrench(cnt_table, condition = group)
  
  # deseq.obj <- DESeqDataSetFromMatrix(cnt_table %>% as.data.frame(), DataFrame(group), ~group)
  # DESeq2::sizeFactors(deseq.obj) <- w$nf
  # cnt_table_normalized <- DESeq2::counts(deseq.obj, normalized=TRUE)
  
  norm_factors <- w$nf
  norm_counts <- sweep(cnt_table, 2, norm_factors, FUN = '/')
  if(roundUp){norm_counts <- norm_counts %>% round()}
  if(!is.null(phy_tree(PhyloObjct, errorIfNULL = F))){
    return(phyloseq(otu_table(norm_counts, taxa_are_rows = T), tax_table(PhyloObjct %>% tax_table()), sample_data(PhyloObjct %>% sample_data()),
                    phy_tree(PhyloObjct)))
  } else{
    return(phyloseq(otu_table(norm_counts, taxa_are_rows = T), tax_table(PhyloObjct %>% tax_table()), sample_data(PhyloObjct %>% sample_data())))
  }
}
##
append_AN_NR <- function(pseq, df_additional){
  df_meta <- meta(pseq)
  if(!"AN_NR" %in%  colnames(df_meta)){
    print("There must be AN_NR column!")
    return(NA)
  }
  ## check if info is already in the main data frame, warning if already exits
  col_names_fea <- setdiff(colnames(df_additional), "AN_NR")
  new_col_names_fea <- setdiff(col_names_fea, colnames(df_meta))
  if(length(new_col_names_fea)==0){
    warning("All new feature already exits, nothing to be added")
    return(pseq)
  }
  if(length(new_col_names_fea) < length(col_names_fea)){
    warning("partially info already there, add the rest!")
    df_additional <- df_additional %>% 
      select(c("AN_NR", new_col_names_fea))
  }
  
  df_meta_new <- df_meta %>% 
    rownames_to_column(var = "SampleName") %>% 
    left_join(., df_additional, by = "AN_NR") %>% 
    column_to_rownames(var = "SampleName")
  sample_data(pseq) <- df_meta_new
  return(pseq)
}
## alpha plot function with LMM with fixed and random variables
AlphaPlot_Violin_LMM <- function(df, SampleID, strata, val, metric = "Obsered species"){
  ## df: in pivot longer manner
  ## calc p value in manner of paired
  ## get y.position for ggplot
  y_pos <- t_test(formula = as.formula(sprintf("%s ~ %s", val, strata)), data = df) %>% add_xy_position() %>% pull(y.position)
  model <- lmer(as.formula(sprintf("%s ~ %s + (1|%s)", val, strata, SampleID)), data = df)
  emm <- emmeans(model, specs = as.formula(sprintf("pairwise ~ %s", strata)))
  contr <- contrast(emm, method = "pairwise")
  
  p_add <- contr %>% 
    as_tibble() %>%
    transmute(
      group1 = sub(" - .*", "", contrast),
      group2 = sub(".* - ", "", contrast),
      estimate,
      se = SE,
      statistic = t.ratio,
      df,
      p = p.value
      ##method = "emmeans model-based t-test"
    ) %>% 
    add_significance() %>% 
    mutate(y.position = y_pos)
  plt <- ggplot(df, aes(x = .data[[strata]], y = .data[[val]], fill = .data[[strata]])) +
    geom_violin() +
    #geom_boxplot(alpha = 0.6, outlier.shape = NA) +
    geom_boxplot(width=0.2) +
    geom_point(position = position_jitter(width = 0.1), alpha = 0.5) +
    # Connect lines for the same sample to show paired changes
    geom_line(aes(group = AN_NR), color = "gray", alpha = 0.5) +
    theme_minimal() +
    labs(
      title = "Comparison of Decontamination Methods",
      subtitle = sprintf("Alpha Diversity: %s", metric),
      y = metric
    ) +
    theme(legend.position = "none")
  plt <- plt + stat_pvalue_manual(p_add, label = "p.signif", inherit.aes = FALSE, tip.length = 0.01)
  return(plt)
}
```

**Load original data**

```{r}
#| include: false
## Load data
#source("../../../add_code/Functions.R")
df_additionalInfo <- readxl::read_xlsx("../../../meta/Mice_meta.xlsx")
pseq_raw <- readRDS("../../../data/Chap3/pseq_Proj5_postFilter_v04.rds") %>% 
  ps_filter(ffpe.bulk == "bulk") %>% 
  ps_filter(true.control == "true") %>% 
  append_AN_NR(df_additional = df_additionalInfo) #%>% 
  #WrenchWrapper(grp = "Sex") 
pseq_restrictive <- readRDS("../../../data/Chap3/pseq_bulk_restrictive.rds") %>% 
  append_AN_NR(df_additionalInfo) #%>% 
  #WrenchWrapper(grp = "Sex")
pseq_decontam <- readRDS("../../../data/Chap3/pseq_bulk_decontam_p0.5.rds") %>% 
  append_AN_NR(df_additionalInfo) #%>% 
  #WrenchWrapper(grp = "Sex")
pseq_SCRuB <- readRDS("../../../data/Chap3/pseq_bulk_SCRuB.rds") %>% 
  append_AN_NR(df_additionalInfo) #%>% 
  #WrenchWrapper(grp = "Sex")
pseq_Nj <- readRDS("../../../data/Chap3/pseq_bulk_Fisher_v02.rds") %>% 
  append_AN_NR(df_additionalInfo) #%>% 
  #WrenchWrapper(grp = "Sex")
```

**Load replica**

```{r}
rep_raw <- readRDS("../../../data/Chap3_Addition/pseq_origin_v1.0.rds") %>% 
  ps_filter(ffpe.bulk == "bulk") %>% 
  append_AN_NR(df_additional = df_additionalInfo)
rep_restrictive <- readRDS("../../../data/Chap3_Addition/pseq_restrictive.rds") %>% 
  append_AN_NR(df_additional = df_additionalInfo)
rep_decontam <- readRDS("../../../data/Chap3_Addition/pseq_decontam.rds") %>% 
  append_AN_NR(df_additional = df_additionalInfo)
rep_SCRuB <- readRDS("../../../data/Chap3_Addition/pseq_SCRuB.rds") %>% 
  append_AN_NR(df_additional = df_additionalInfo)
rep_Nj <- readRDS("../../../data/Chap3_Addition/pseq_Nejman.rds") %>% 
  append_AN_NR(df_additional = df_additionalInfo)
## rename taxa
taxa_names(rep_raw) <- gsub(pattern = " ", replacement = "_", x = taxa_names(rep_raw))
taxa_names(rep_restrictive) <- gsub(pattern = " ", replacement = "_", x = taxa_names(rep_restrictive))
taxa_names(rep_decontam) <- gsub(pattern = " ", replacement = "_", x = taxa_names(rep_decontam))
taxa_names(rep_SCRuB) <- gsub(pattern = " ", replacement = "_", x = taxa_names(rep_SCRuB))
taxa_names(rep_Nj) <- gsub(pattern = " ", replacement = "_", x = taxa_names(rep_Nj))
```

Retain only sample with pair

```{r}
IDs <- meta(rep_raw)[["AN_NR"]]
pseq_raw <- pseq_raw %>% ps_filter(AN_NR %in% IDs)
pseq_restrictive <- pseq_restrictive %>% ps_filter(AN_NR %in% IDs)
pseq_decontam <- pseq_decontam %>% ps_filter(AN_NR %in% IDs)
pseq_SCRuB <- pseq_SCRuB %>% ps_filter(AN_NR %in% IDs)
pseq_Nj <- pseq_Nj %>% ps_filter(AN_NR %in% IDs)
```

# Project Structure through Mind-Map

```{mermaid}
flowchart TD
  %%A[Hard edge] --> B(Round edge)
  %%A --> F(Others)
  %%B --> C{Decision}
  %%C --> D[Result one]
  %%C --> E[Result two]
  
  S(Replica Analysis) --> B(Properties of Replica)
  S --> C(Overlap)
  S --> D(Assessment)
  
```

# Replica Properties

## Alpha

Diversity between

## Beta

# Assessment

Using replica as additional dataset for assessment of decontamination approaches.

## Intra-samples Stability

-   **Assumption**: after removing contaminants, bacterial profile of a sample should be stable, regardless of sequencing year or technician.
-   **Method**: using appropriate distance function to measure the (dis)similarity of the sample sample w.r.t 2 technical replicas. Afterwards, applying paired t-test or Wilcoxon Signed-Rank Test, LMM, or constrained permutation test.

### Bray-curtis distance with compositional normalization

```{r}
## Write a wrapper

process_replica <- function(
    ps1, ## original sample
    ps2, ## replica 
    normalize = "compositional", ## compositional - bray | clr - euclidean
    distance = "bray",
    col_name = "raw"){
  ## add replica info and change names
  ps1 <- ps1 %>% 
    ps_mutate(Replica = "R1") %>% 
    ps_mutate(NewName = sprintf("%s_%s", AN_NR, Replica))
  sample_names(ps1) <- meta(ps1)[["NewName"]]
  
  ps2 <- ps2 %>% 
    ps_mutate(Replica = "R2") %>% 
    ps_mutate(NewName = sprintf("%s_%s", AN_NR, Replica))
  sample_names(ps2) <- meta(ps2)[["NewName"]]
  
  dis_mx <- merge_phyloseq(ps1, ps2) %>% 
    microbiome::transform(transform = normalize) %>% 
    phyloseq::distance(method = distance) %>% 
    as.matrix() %>% 
    as.data.frame() %>% 
    dplyr::select(contains("R2")) %>% 
    rownames_to_column(var = "rowname") %>% 
    filter(grepl("R1", rowname)) %>% 
    arrange(rowname) %>% 
    column_to_rownames("rowname") %>% 
    select(sort(names(.)))
  
  col <- tibble(AN_NR = rownames(dis_mx), !!sym(col_name) := diag(as.matrix(dis_mx)))
  return(col)
}

# pseq_raw <- pseq_raw %>% 
#   ps_mutate(Replica = "R1") %>% 
#   ps_mutate(NewName = sprintf("%s_%s", AN_NR, Replica))
# sample_names(pseq_raw) <- meta(pseq_raw)[["NewName"]]
# 
# rep_raw <- rep_raw %>% 
#   ps_mutate(Replica = "R2") %>% 
#   ps_mutate(NewName = sprintf("%s_%s", AN_NR, Replica))
# sample_names(rep_raw) <- meta(rep_raw)[["NewName"]]
# 
# merge_raw <- merge_phyloseq(pseq_raw, rep_raw)
# 
# dis_mx <- merge_raw %>% microbiome::transform(transform = "compositional") %>% phyloseq::distance(method = "bray") %>% as.matrix()
# xx <- dis_mx %>% 
#   as.data.frame() %>% 
#   dplyr::select(contains("R2")) %>% 
#   rownames_to_column(var = "rowname") %>% 
#   filter(grepl("R1", rowname)) %>% 
#   arrange(rowname) %>% 
#   column_to_rownames("rowname") %>% 
#   select(sort(names(.)))
# 
# tmp <- merge_phyloseq(pseq_raw, rep_raw) %>% 
#     microbiome::transform(transform = "compositional") %>% 
#     phyloseq::distance(method = "bray") %>% 
#     as.data.frame() %>% 
#     dplyr::select(contains("R2")) %>% 
#     rownames_to_column(var = "rowname") %>% 
#     filter(grepl("R1", rowname)) %>% 
#     arrange(rowname) %>% 
#     column_to_rownames("rowname") %>% 
#     select(sort(names(.)))
# 
# col <- tibble(AN_NR = rownames(xx), raw = diag(as.matrix(xx)))

## create a wide format table
## AN_NR  Raw Restrictive Decontam  SCRuB Nj
## 1    dist(.,.) dist(.,.) ## distance between 2 replica
```

Create wide-format table

```{r}
df <- process_replica(pseq_raw, rep_raw, col_name = "Raw") %>% 
  merge(., process_replica(pseq_restrictive, rep_restrictive, col_name = "Restrictive"), by = "AN_NR") %>%
  merge(., process_replica(pseq_decontam, rep_decontam, col_name = "Decontam"), by="AN_NR") %>% 
  merge(., process_replica(pseq_SCRuB, rep_SCRuB, col_name = "SCRuB"), by="AN_NR") %>% 
  merge(., process_replica(pseq_Nj, rep_Nj, col_name = "Nj"), by="AN_NR")
df_longer <- df %>% 
  pivot_longer(cols = -AN_NR, names_to = "Methods", values_to = "Val")
```

Alpha

```{r}
AlphaPlot_Violin_LMM(df_longer, "AN_NR", "Methods", "Val")
```
